### 2010-09-21 docs/CrystalPlan_paper/paper.tex
Refs #182: Template for a latex CrystalPlan paper.

%------------------------------------------------------------------------------
% Template file for the submission of papers to IUCr journals in LaTeX2e
% using the iucr document class
% Copyright 1999-2009 International Union of Crystallography
% Version 1.4 (11 May 2009)
%------------------------------------------------------------------------------

\documentclass{iucr}              % DO NOT DELETE THIS LINE

     %-------------------------------------------------------------------------
     % Information about the type of paper
     %-------------------------------------------------------------------------
     \paperprodcode{a000000}      % Replace with production code if known
     \paperref{xx9999}            % Replace xx9999 with reference code if known
     \papertype{FA}               % Indicate type of article
                                  %   FA - research papers (full article)
                                  %   SC - short communications
                                  %   LA - lead article
                                  %   FE - feature articles
                                  %   ST - structural communications
                                  %   XC - crystallization communications
                                  % (Following categories rarely in LaTeX)
                                  %   AA - abstracts
                                  %   AD - addenda and errata
                                  %   BC - books received
                                  %   BR - book reviews
                                  %   CA - cif applications
                                  %   CE - current events
                                  %   CI - inorganic compounds
                                  %   CM - metal-organic compounds
                                  %   CN - cryocrystallography papers
                                  %   CO - organic compounds
                                  %   CP - computer programs
                                  %   CR - crystallographers
                                  %   CS - scientific comment
                                  %   ED - editorial
                                  %   EI - inorganic compounds
                                  %   EM - metal-organic compounds
                                  %   EO - organic compounds
                                  %   FI - inorganic compounds
                                  %   FM - metal-organic compounds
                                  %   FO - organic compounds
                                  %   IP - issue preface
                                  %   IU - iucr
                                  %   LE - letters to the editor
                                  %   LN - laboratory notes
                                  %   ME - forthcoming meetings/short courses
                                  %   MR - meeting reports
                                  %   NN - notes and news
                                  %   NP - new commercial products
                                  %   OB - obituaries
                                  %   SR - software reviews
                                  %   TE - teaching and education

     \paperlang{english}          % Can be english, french, german or russian
     %-------------------------------------------------------------------------
     % Information about journal to which submitted
     %-------------------------------------------------------------------------
     \journalcode{J}              % Indicate the journal to which submitted
                                  %   A - Acta Crystallographica Section A
                                  %   B - Acta Crystallographica Section B
                                  %   C - Acta Crystallographica Section C
                                  %   D - Acta Crystallographica Section D
                                  %   E - Acta Crystallographica Section E
                                  %   F - Acta Crystallographica Section F
                                  %   J - Journal of Applied Crystallography
                                  %   S - Journal of Synchrotron Radiation
          %--------------------------------------------------------------------
          % The following entries will be changed as required by editorial staff
          %--------------------------------------------------------------------
     \journalyr{2010}
     \journaliss{1}
     \journalvol{65}
     \journalfirstpage{000}
     \journallastpage{000}
     \journalreceived{0 XXXXXXX 0000}
     \journalaccepted{0 XXXXXXX 0000}
     \journalonline{0 XXXXXXX 0000}

\begin{document}                  % DO NOT DELETE THIS LINE

     %-------------------------------------------------------------------------
     % The introductory (header) part of the paper
     %-------------------------------------------------------------------------

     % The title of the paper. Use \shorttitle to indicate an abbreviated title
     % for use in running heads (you will need to uncomment it).

\title{CrystalPlan: an Experiment Planning Tool for Crystallography}
\shorttitle{CrystalPlan, an Experiment Planning Tool}

     % Authors' names and addresses. Use \cauthor for the main (contact) author.
     % Use \author for all other authors. Use \aff for authors' affiliations.
     % Use lower-case letters in square brackets to link authors to their
     % affiliations; if there is only one affiliation address, remove the [a].

\cauthor[a]{Janik}{Zikovsky}{zikovskyjl@ornl.gov}{address if different from
\aff}
\author[a]{Peter}{Peterson}
\author[a]{Xiaoping}{Wang}
\author[a]{Matthew}{Frost}
\author[a]{Christina}{Hoffmann}

\aff[a]{Spallation Neutron Source, Oak Ridge National Laboratory, P.O. Box 2008
MS-6477, Oak Ridge, TN 37831-6477
\country{USA}}

     % Use \shortauthor to indicate an abbreviated author list for use in
     % running heads (you will need to uncomment it).

\shortauthor{Zikovsky, Peterson, Wang, Frost and Hoffmann}

     % Use \vita if required to give biographical details (for authors of
     % invited review papers only). Uncomment it.

%\vita{Author's biography}

     % Keywords (required for Journal of Synchrotron Radiation only)
     % Use the \keyword macro for each word or phrase, e.g. 
     % \keyword{X-ray diffraction}\keyword{muscle}

%\keyword{keyword}

     % PDB and NDB reference codes for structures referenced in the article and
     % deposited with the Protein Data Bank and Nucleic Acids Database (Acta
     % Crystallographica Section D). Repeat for each separate structure e.g
     % \PDBref[dethiobiotin synthetase]{1byi} \NDBref[d(G$_4$CGC$_4$)]{ad0002}

%\PDBref[optional name]{refcode}
%\NDBref[optional name]{refcode}

\maketitle                        % DO NOT DELETE THIS LINE

\begin{synopsis}
Supply a synopsis of the paper for inclusion in the Table of Contents.
\end{synopsis}

\begin{abstract}
Beam time at large x-ray and neutron scattering facilities is always at a premium.
The CrystalPlan program can calculate the data coverage of a crystal in reciprocal
space in a single-crystal diffraction time-of-flight experiment. CrystalPlan can 
help a user build an experiment plan that will acquire the most data possible, 
with sufficient coverage but limited redundancy, therefore increasing scientific 
productivity. 
An attractive GUI including a 3D viewer and an automated coverage optimizer 
are among its useful features. 
A sample use case of the program with the TOPAZ beamline at SNS will be
presented. 
\end{abstract}

     %-------------------------------------------------------------------------
     % The main body of the paper
     %-------------------------------------------------------------------------
     % Now enter the text of the document in multiple \section's, \subsection's
     % and \subsubsection's as required.

\section{Section title}

Text text text text text text text text text text text text text text
text text text text text text text.

\subsection{Title}

Text text text text text text text text text text text text text text
text text text text text text text.

\subsubsection{Title}

Text text text text text text text text text text text text text text
text text text text text text text.

     % Appendices appear after the main body of the text. They are prefixed by
     % a single \appendix declaration, and are then structured just like the
     % body text.

\appendix
\section{Appendix title}

Text text text text text text text text text text text text text text
text text text text text text text.

\subsection{Title}

Text text text text text text text text text text text text text text
text text text text text text text.

\subsubsection{Title}

Text text text text text text text text text text text text text text
text text text text text text text.

     %-------------------------------------------------------------------------
     % The back matter of the paper - acknowledgements and references
     %-------------------------------------------------------------------------

     % Acknowledgements come after the appendices

\ack{Acknowledgements}

     % References are at the end of the document, between \begin{references}
     % and \end{references} tags. Each reference is in a \reference entry.

\begin{references}
\reference{Author, A. \& Author, B. (1984). \emph{Journal} \textbf{Vol}, 
first page--last page.}
\end{references}

     %-------------------------------------------------------------------------
     % TABLES AND FIGURES SHOULD BE INSERTED AFTER THE MAIN BODY OF THE TEXT
     %-------------------------------------------------------------------------

     % Simple tables should use the tabular environment according to this
     % model

\begin{table}
\caption{Caption to table}
\begin{tabular}{llcr}      % Alignment for each cell: l=left, c=center, r=right
 HEADING    & FOR        & EACH       & COLUMN     \\
\hline
 entry      & entry      & entry      & entry      \\
 entry      & entry      & entry      & entry      \\
 entry      & entry      & entry      & entry      \\
\end{tabular}
\end{table}

     % Postscript figures can be included with multiple figure blocks

\begin{figure}
\caption{Caption describing figure.}
\includegraphics{fig1.ps}
\end{figure}

\end{document}                    % DO NOT DELETE THIS LINE
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2010-09-21 docs/CrystalPlan_paper/paper.tex
Refs #182: copied bulk of the Word document into the latex file.

@@
 \title{CrystalPlan: an Experiment Planning Tool for Crystallography}
-\shorttitle{CrystalPlan, an Experiment Planning Tool}
+\shorttitle{CrystalPlan: an Experiment Planning Tool}
 
@@
 
-\shortauthor{Zikovsky, Peterson, Wang, Frost and Hoffmann}
-
-     % Use \vita if required to give biographical details (for authors of
-     % invited review papers only). Uncomment it.
-
-%\vita{Author's biography}
+\shortauthor{Zikovsky et al.}
 
@@
 
-     % PDB and NDB reference codes for structures referenced in the article and
-     % deposited with the Protein Data Bank and Nucleic Acids Database (Acta
-     % Crystallographica Section D). Repeat for each separate structure e.g
-     % \PDBref[dethiobiotin synthetase]{1byi} \NDBref[d(G$_4$CGC$_4$)]{ad0002}
-
-%\PDBref[optional name]{refcode}
-%\NDBref[optional name]{refcode}
-
 \maketitle                        % DO NOT DELETE THIS LINE
@@
 \begin{synopsis}
-Supply a synopsis of the paper for inclusion in the Table of Contents.
+Describes CrystalPlan, a software program for the automatic creation of 
+crystallography experiment plans for time-of-flight diffractometers.
 \end{synopsis}
@@
 %-------------------------------------------------------------------------
-     % The main body of the paper
 %-------------------------------------------------------------------------
-     % Now enter the text of the document in multiple \section's, \subsection's
-     % and \subsubsection's as required.
+%---------------------------------------------------------------------------------------
+\section{Introduction}
 
-\section{Section title}
+The TOPAZ beamline at the Spallation Neutron Source (SNS) at Oak Ridge National Lab (ORNL),
+currently finishing commissioning, is a neutron time-of-flight instrument designed to 
+acquire single-crystal diffraction data of small-to-moderate sized unit cells with high throughput. 
+The final design calls for an array of 48 detectors covering a large fraction of
+$2\pi$ steradians; however as of this writing, only 14 of the 48 detectors are
+installed. In order to have sufficient data to perform a fit, an experimenter will
+typically want to measure $ > 85\%$ of the peaks within a certain
+range of d-spacings. However, TOPAZ's complex geometry makes it very difficult to 
+come up with an effective set of sample orientations that will
+cover all the required peaks without unnecessary redundancy.
 
-Text text text text text text text text text text text text text text
-text text text text text text text.
+Existing software tools [ADD SOME NAMES HERE – see my notes from ACA2010] 
+were designed for X-ray experiments where a single incident
+wavelength and a large number of sample orientations or sweeps are used to
+collect data. These programs cannot be used for TOPAZ, 
+since as a time-of-flight instrument, it is able to collect data
+from a wide bandwidth of incident neutrons. 
+Additionally, this means that relatively few sample orientations are needed in
+order to complete a measurement: whereas an X-ray measurement might have
+hundreds or thousands of individual frames, a TOPAZ run may have as few as 10.
 
-\subsection{Title}
+The goal of CrystalPlan was to make an easy-to-use software tool to aid users in
+planning their experiments. By quickly making an experimental plan that will be
+sufficient for the needs of the scientist (enough peaks measured), but without
+spending time needlessly measuring the same region in reciprocal space,
+CrystalPlan will greatly increase the scientific productivity and throughput of
+TOPAZ. More users will be able to measure more samples, faster.
 
-Text text text text text text text text text text text text text text
-text text text text text text text.
+%---------------------------------------------------------------------------------------
+%---------------------------------------------------------------------------------------
+%---------------------------------------------------------------------------------------
+\section{Architecture}
 
-\subsubsection{Title}
+The program needs to be cross-platform, and run equally well on Linux, Mac OS
+and Windows, although Linux is our primary target operating system. Another
+design goal was to use of open-source software and toolkits as much as possible.
 
-Text text text text text text text text text text text text text text
-text text text text text text text.
+The calculation code of CrystalPlan is placed in a model layer and kept
+completely separate from eventual GUI elements. This makes it possible to run
+calculations using Python scripts, if desired. Unit tests for each part of the
+calculations are used to ensure consistent results on different platforms and
+systems.
+
+\subsection{Instrument Configuration}
+
+CrystalPlan supports area detectors with flat faces and arbitrary rectangular
+dimensions; this handles the detector types in use at SNS (Anger cameras and
+8-pack tube detectors), but the software could easily be extended for more
+complicated detector shapes as needed. The detector configuration is loaded from
+a text file containing the location, size, orientation and number of pixels of
+each detector.
+
+\subsection{Sample Orientation}
+
+CrystalPlan is designed to be able to use a variety of sample orientation
+goniometers. A base Goniometer class is subclassed for specific types of
+goniometers, allowing a unified API: in a way, it is a “universal goniometer”,
+which uses the conventional goniometer angles of phi, chi, omega as defined in
+[REF XXX]. Each class inheriting from Goniometer can convert the desired phi,
+chi, omega angles to the internal motor positions it requires. The API allows
+for limited goniometers: for example, TOPAZ is currently equipped with a
+goniometer where chi is fixed at 45 degrees. CrystalPlan handles this by
+limiting the available sample orientation angles accessible to the user.
+
+The data analysis program ISAW is used to calculate the the sample mounting
+orientation UB matrix, which is saved to a text file along with the lattice
+parameters. CrystalPlan can load this UB matrix file and will use that UB matrix
+in coverage calculations.
+
+\subsection{Volume coverage}
+
+\subsubsection{Reciprocal Space Representation}
+
+The reciprocal space to be measured is divided into a cubic 3D matrix, the size
+of which is determined by the user, who specifies the minimum d-spacing
+($d_{min}$, in $\AA$) that he or she is interested in. The resolution of the 3D
+matrix is also specified (in $\AA^{-1}$), which determines the number of points to be
+modeled. Machine memory limits the resolution that can be achieved.
+
+Each point of the reciprocal space 3D matrix holds an integer representing the 
+number of times that voxel of q-space has been measured – this is the total
+coverage matrix. To speed up the calculation of the total, a coverage matrix for
+each sample orientation is calculated and saved in memory. This means that the
+total coverage matrix can be computed by simply adding multiple 3D matrices
+corresponding to each sample orientation.
+
+\subsubsection{Calculating Coverage of One Sample Orientation}
+
+For each sample orientation, a rotation matrix combining the sample's mounting
+orientation (the U matrix loaded from ISAW) and the goniometer rotation
+($\Phi$, $\Xi$, $\Omega$) is computed. The coverage in reciprocal space is
+computed in this way: for each pixel in the face of a detector, the direction of
+the scattered beam is known. Then, two incident beam vectors are considered,
+corresponding to the low and high wavelength limits of the detectors and/or
+source. The scattered beam direction and the two possible incident beam vectors
+are used to to find two q vectors, which point in the same direction. At the
+high wavelength we find qmin, and the low wavelength limit scattering off of qmax.
+Since the instrument has a wide bandwidth, we fill in the 3D q-space matrix by
+filling all points from qmin to the  qmax, for all pixels in the face of the
+detector. This quickly defines a volume of coverage, as shown in Figure X.
+
+To reduce memory usage yet allow for quickly trying different detector
+configurations, the 3D matrix of each sample orientation is saved a 32- or
+64-bit integer, with each bit representing a particular detector. A simple
+binary mask can then be applied to quickly compute the coverage if some
+detectors are disabled, for example.
+
+\subsection{Crystal Parameters and Single Reflection Coverage}
+
+Crystal lattice parameters (lengths a,b,c and angles alpha, beta, gamma) are
+typed in or loaded from the ISAW UB matrix file. CrystalPlan then generates a
+list of HKL peaks with d-spacings larger than the $d_{min}$ specified. For each
+sample orientation being simulated and for each peak, a function calculates at
+what wavelength lambda and in which direction it scatters. If the wavelength is
+within range of the limits of the instrument, the beam direction is projected
+onto each detector face to determine where, on each detector surface, it will be
+measured.
+
+Each HKL peak is a separate object that holds a list of simulated measurements,
+including which detector measured it, at what position on the detector face, at
+what wavelength, and for which sample orientation. These data will be used to
+graphically display coverage and calculate statistics such as redundancy.
+
+\subsection{Crystal Symmetry}
+
+When entering the lattice parameters, the crystal's point group symmetry can
+also be entered. This information is then used to calculate the increased
+measurement redundancy. A list of n 3x3 multiplication matrices is generated
+based on the point group; any {hkl} vector can be multiplied by these matrices
+to find the n equivalent hkl reflections; for example, for orthorombic (mmm)
+symmetry, n=8 and 8 multiplication matrices are generated, which multiply {hkl}
+to give {±h, ±k, ±l}.
+
+For the single reflections, each hkl is tracked separately, but a primary hkl is
+found that points to a list of equivalent hkls. The GUI can then display either
+all hkl, ignoring crystal symmetry; or combine them to display only part of the
+reflections.
+
+In volume coverage mode, a 4D matrix is generated where each voxel has a list of
+n corresponding equivalent voxels. This allows CrystalPlan to quickly compute a
+volume coverage map that takes into account crystal symmetry: at each voxel, n
+original voxels are checked to see if any of them were measured.
+
+\subsection{Automatic Coverage Optimizer}
+
+CrystalPlan allows the user to manually enter lists of sample orientations to
+simulate; however, finding an optimal plan by trial and error would be very time
+consuming. To make the process easier, CrystalPlan includes an Automatic
+Coverage Optimizer, which takes as its inputs the desired number of sample
+orientations and uses a genetic algorithm to find a solution that satisfies a
+user-specified coverage criterion (for example, the user may ask for 85% of
+peaks to be measured, or for 90% of q-volume).
+
+Genetic algorithms have been covered in the literature before, see e.g ref
+[INSERT REFERENCE HERE]. In this application, the genes consist of goniometer
+angles. For instance, if a goniometer has freedom in 3 angles and the user
+requests a 10 different sample orientations, then each individual has 30 genes.
+The initial population is created randomly and is made to evolve using mutations
+and crossover. For each individual set of sample orientations, the coverage
+statistics are computed (taking crystal symmetry into account if desired), and
+this is used as the fitness of the given individual in the normal genetic
+algorithm process.
+
+The Automatic Coverage Optimizer can find an acceptable solution (if any is
+possible) in times ranging from a few seconds to several minutes, depending on
+the difficulty of the problem and the computing hardware. Given that at a
+neutron facility, a measurement at a single sample orientation may take a few
+hours, this feature can result in significant beam-time savings.
+
+%---------------------------------------------------------------------------------------
+%---------------------------------------------------------------------------------------
+%---------------------------------------------------------------------------------------
+\section{Implementation}
+
+\subsection{Language and Libraries}
+
+CrystalPlan was written in Python, a powerful, open, cross-platform scripting
+language. Python was chosen for its ease of cross-platform deployment, and for
+its attractive language features. The numpy and scipy libraries (two
+open-source, commonly installed Python libraries) were used for many of the
+calculations. Some of the critical calculations were also written with inline C
+code, which speed up some calculations by up to $400\times$ as compared with
+the pure Python version (which can still be used, in case of incompatibilities).
+
+\subsection{Graphical User Interface}
+
+\subsubsection{Design}
+
+An attractive and easy-to-use GUI was considered an important part of
+CrystalPlan, since the program is meant to be used interactively as a guide to
+the experimenter. The user interface was written using wxPython, a mature
+cross-platform GUI toolkit, as well as the Enthought Traits GUI. Some of the 2D
+plots use matplotlib, and the 3D visualization elements use the Mayavi2 toolkit
+for Python, itself based on VTK. All of these libraries are open-source.
+
+\subsubsection{Main Program Workflow}
+Some screenshots of the GUI are shown in Figure Y. The main window consists of a
+series of tabs, with workflow proceeding from left to right:
+
+\begin{itemize}
+  \item Q-Space tab: define the volume of interest (minimum d-spacing) and
+resolution of reciprocal space. Detectors tab: load instrument detector geometry
+files, and enable or disable individual detectors. Goniometer tab: to choose the
+goniometer and degrees of orientation freedom of the instrument.
+
+  \item Sample tab: enter or load the crystal's lattice parameters and UB matrix.
+
+  \item Try an Orientation tab: interactively rotate a sample and observe the
+  changes in coverage in real-time.
+
+  \item Add Orientations tab: type in a list of sample orientation angles to add
+  to the experiment plan.
+
+  \item Experiment Plan tab: a list of the previously calculated sample orientations.
+Here you can manage the orientations you will use in the experiment: add,
+delete, or temporarily disable them to see the effect on coverage. Once
+complete, the plan can be sent to the SNS data acquisition computers in order to
+begin a run.
+
+\end{itemize}
+
+\subsubsection{Reciprocal Space Coverage Visualization}
+Figures Z -a shows a screenshot of the reciprocal space coverage 3D interface.
+The 3D view can represent q-space in two ways: volume coverage, or single-peak
+coverage.
+
+The volume coverage shows a solid isosurface representing the volume measured at
+least once. Redundant regions can also be shown using semi-transparent
+isosurfaces; or the coverage can be inverted, showing non-measured volumes only;
+or sliced between user-selectable $q_{min}$ and $q_{max}$.
+
+The reflections view (Figure ZZ-a) displays each hkl as a small sphere; the
+color of which indicates the number of times the hkl is predicted to be measured
+in the current experiment plan.
+
+Both views can take into crystal symmetry in their display (Figures Z-b and
+ZZ-b), using the techniques described previously.
+
+Finally, both views display quick coverage statistics, showing the percentage of
+volume (or of the number of peaks) measured, as well as the proportion of
+redundancy (voxels or peaks measured 2 or more times).
+
+\section{Sample Use Case}
+
+\section{Conclusions and Future Development}
+
+Live data processing    …  interface with ISAW … InelasCasetic instruments ??
 
@@
 
-\appendix
-\section{Appendix title}
+%\appendix
+%\section{Appendix title}
 
-Text text text text text text text text text text text text text text
-text text text text text text text.
-
-\subsection{Title}
-
-Text text text text text text text text text text text text text text
-text text text text text text text.
-
-\subsubsection{Title}
-
-Text text text text text text text text text text text text text text
-text text text text text text text.
+%Text text text text text text text text text text text text text text
+%text text text text text text text.
 
@@
 \ack{Acknowledgements}
+
+This research is supported by UT Battelle, LLC under Contract No.
+DE-AC05-00OR22725 for the U.S. Department of Energy, Office of Science.
 
@@
 \caption{Caption describing figure.}
-\includegraphics{fig1.ps}
+%\includegraphics{fig1.ps}
 \end{figure}

### 2010-09-21 docs/CrystalPlan_paper/paper.tex
Refs #182: bibliography and figures set up.

@@
                                   %   XC - crystallization communications
-                                  % (Following categories rarely in LaTeX)
-                                  %   AA - abstracts
-                                  %   AD - addenda and errata
-                                  %   BC - books received
-                                  %   BR - book reviews
-                                  %   CA - cif applications
-                                  %   CE - current events
-                                  %   CI - inorganic compounds
-                                  %   CM - metal-organic compounds
-                                  %   CN - cryocrystallography papers
-                                  %   CO - organic compounds
-                                  %   CP - computer programs
-                                  %   CR - crystallographers
-                                  %   CS - scientific comment
-                                  %   ED - editorial
-                                  %   EI - inorganic compounds
-                                  %   EM - metal-organic compounds
-                                  %   EO - organic compounds
-                                  %   FI - inorganic compounds
-                                  %   FM - metal-organic compounds
-                                  %   FO - organic compounds
-                                  %   IP - issue preface
-                                  %   IU - iucr
-                                  %   LE - letters to the editor
-                                  %   LN - laboratory notes
-                                  %   ME - forthcoming meetings/short courses
-                                  %   MR - meeting reports
-                                  %   NN - notes and news
-                                  %   NP - new commercial products
-                                  %   OB - obituaries
-                                  %   SR - software reviews
-                                  %   TE - teaching and education
 
@@
 
-     % The title of the paper. Use \shorttitle to indicate an abbreviated title
-     % for use in running heads (you will need to uncomment it).
+\title{CrystalPlan: an Experiment Planning Tool for Crystallography}
 
-\title{CrystalPlan: an Experiment Planning Tool for Crystallography}
 \shorttitle{CrystalPlan: an Experiment Planning Tool}
@@
 
-     % Keywords (required for Journal of Synchrotron Radiation only)
-     % Use the \keyword macro for each word or phrase, e.g. 
-     % \keyword{X-ray diffraction}\keyword{muscle}
-
-%\keyword{keyword}
-
 \maketitle                        % DO NOT DELETE THIS LINE
@@
 %-------------------------------------------------------------------------
-%---------------------------------------------------------------------------------------
+%-------------------------------------------------------------------------
 \section{Introduction}
 
-The TOPAZ beamline at the Spallation Neutron Source (SNS) at Oak Ridge National Lab (ORNL),
-currently finishing commissioning, is a neutron time-of-flight instrument designed to 
-acquire single-crystal diffraction data of small-to-moderate sized unit cells with high throughput. 
-The final design calls for an array of 48 detectors covering a large fraction of
-$2\pi$ steradians; however as of this writing, only 14 of the 48 detectors are
-installed. In order to have sufficient data to perform a fit, an experimenter will
-typically want to measure $ > 85\%$ of the peaks within a certain
-range of d-spacings. However, TOPAZ's complex geometry makes it very difficult to 
-come up with an effective set of sample orientations that will
-cover all the required peaks without unnecessary redundancy.
+The TOPAZ beamline at the Spallation Neutron Source (SNS) at Oak Ridge National
+Lab (ORNL), currently finishing commissioning, is a neutron time-of-flight
+instrument designed to acquire single-crystal diffraction data of
+small-to-moderate sized unit cells with high throughput. The final design calls
+for an array of 48 detectors covering a large fraction of  $2\pi$ steradians;
+however as of this writing, only 14 of the 48 detectors are installed. In order
+to have sufficient data to perform a fit, an experimenter will  typically want
+to measure $ > 85\%$ of the peaks within a certain range of d-spacings. 
+However, TOPAZ's complex geometry makes it very difficult to come up with an
+effective set of sample orientations that will cover all the required peaks
+without unnecessary redundancy.
 
@@
 goniometers, allowing a unified API: in a way, it is a “universal goniometer”,
-which uses the conventional goniometer angles of phi, chi, omega as defined in
-[REF XXX]. Each class inheriting from Goniometer can convert the desired phi,
-chi, omega angles to the internal motor positions it requires. The API allows
+which uses the conventional goniometer angles of $\phi$, $\chi$, $\omega$ as
+defined in \cite{knuth84}. Each class inheriting from Goniometer can convert
+the desired phi, chi, omega angles to the internal motor positions it requires. The API allows
 for limited goniometers: for example, TOPAZ is currently equipped with a
@@
 filling all points from qmin to the  qmax, for all pixels in the face of the
-detector. This quickly defines a volume of coverage, as shown in Figure X.
+detector. This quickly defines a volume of coverage, as shown in Figure
+\ref{fig:volume_coverage}.
 
@@
 
-Crystal lattice parameters (lengths a,b,c and angles alpha, beta, gamma) are
-typed in or loaded from the ISAW UB matrix file. CrystalPlan then generates a
-list of HKL peaks with d-spacings larger than the $d_{min}$ specified. For each
+Crystal lattice parameters (lengths $a$, $b$, $c$ and angles $\alpha$, $\beta$,
+$\gamma$) are typed in or loaded from the ISAW UB matrix file. CrystalPlan then
+generates a list of HKL peaks with d-spacings larger than the $d_{min}$ specified. For each
 sample orientation being simulated and for each peak, a function calculates at
@@
 also be entered. This information is then used to calculate the increased
-measurement redundancy. A list of n 3x3 multiplication matrices is generated
-based on the point group; any {hkl} vector can be multiplied by these matrices
+measurement redundancy. A list of $n$ $3\times3$ multiplication matrices is
+generated based on the point group; any {hkl} vector can be multiplied by these matrices
 to find the n equivalent hkl reflections; for example, for orthorombic (mmm)
-symmetry, n=8 and 8 multiplication matrices are generated, which multiply {hkl}
-to give {±h, ±k, ±l}.
+symmetry, $n=8$ and 8 multiplication matrices are generated, which multiply
+{hkl} to give {±h, ±k, ±l}.
 
@@
 orientations and uses a genetic algorithm to find a solution that satisfies a
-user-specified coverage criterion (for example, the user may ask for 85% of
-peaks to be measured, or for 90% of q-volume).
+user-specified coverage criterion (for example, the user may ask for $85\%$ of
+peaks to be measured, or for $90\%$ of q-volume).
 
@@
 \subsubsection{Main Program Workflow}
-Some screenshots of the GUI are shown in Figure Y. The main window consists of a
-series of tabs, with workflow proceeding from left to right:
+Some screenshots of the GUI are shown in Figure \ref{fig:screenshots}. The main
+window consists of a series of tabs, with workflow proceeding from left to right:
 
@@
 \subsubsection{Reciprocal Space Coverage Visualization}
-Figures Z -a shows a screenshot of the reciprocal space coverage 3D interface.
-The 3D view can represent q-space in two ways: volume coverage, or single-peak
+Figure \ref{fig:volume_view}a) shows a screenshot of the reciprocal space
+coverage 3D interface. The 3D view can represent q-space in two ways: volume coverage, or single-peak
 coverage.
@@
 
-The reflections view (Figure ZZ-a) displays each hkl as a small sphere; the
-color of which indicates the number of times the hkl is predicted to be measured
+The reflections view (Figure \ref{fig:refl_view}a) displays each hkl as a
+small sphere; the color of which indicates the number of times the hkl is predicted to be measured
 in the current experiment plan.
 
-Both views can take into crystal symmetry in their display (Figures Z-b and
-ZZ-b), using the techniques described previously.
+Both views can take into crystal symmetry in their display (Figures
+\ref{fig:volume_view}b and \ref{fig:refl_view}b), using the techniques described
+previously.
 
@@
      % and \end{references} tags. Each reference is in a \reference entry.
+\bibliographystyle{iucr}     
+\bibliography{crystalplan}
 
-\begin{references}
-\reference{Author, A. \& Author, B. (1984). \emph{Journal} \textbf{Vol}, 
-first page--last page.}
-\end{references}
+%\begin{references}
+%\reference{Author, A. \& Author, B. (1984). \emph{Journal} \textbf{Vol}, 
+%first page--last page.}
+%\end{references}
 
@@
 \begin{figure}
-\caption{Caption describing figure.}
+\caption{volume coverage}
 %\includegraphics{fig1.ps}
+\label{fig:volume_coverage}
+\end{figure}
+
+\
+\begin{figure}
+\caption{Screenshots of the main GUI}
+%\includegraphics{fig1.ps}
+\label{fig:screenshots}
+\end{figure}
+
+\begin{figure}
+\caption{Volume coverage in reciprocal space view: a) without crystal symmetry;
+b) with crystal symmetry}
+%\includegraphics{fig1.ps}
+\label{fig:volume_view}
+\end{figure}
+
+\begin{figure}
+\caption{Coverage of single-crystal reflections in reciprocal space: a) without
+crystal symmetry; b) with crystal symmetry}
+%\includegraphics{fig1.ps}
+\label{fig:refl_view}
+\end{figure}
+
+\begin{figure}
+\caption{figureeeeee}
+%\includegraphics{fig1.ps}
+\label{fig:screenshots}
 \end{figure}

### 2010-09-21 docs/CrystalPlan_paper/paper.tex
Refs #182: bibliography and figures set up.

@@
 
-Existing software tools [ADD SOME NAMES HERE – see my notes from ACA2010] 
-were designed for X-ray experiments where a single incident
-wavelength and a large number of sample orientations or sweeps are used to
-collect data. These programs cannot be used for TOPAZ, 
-since as a time-of-flight instrument, it is able to collect data
-from a wide bandwidth of incident neutrons. 
+Existing diffraction experiment planning software programs (for example,
+CrysAlis Pro from Oxford Diffraction; DIFFRAC from Bruker AXS) were designed for
+X-ray experiments where a single  incident wavelength and a large number of
+sample orientations or sweeps are used to collect data. 
+These programs cannot be used for TOPAZ, since as a time-of-flight instrument, 
+it is able to collect data from a wide bandwidth of incident neutrons. 
 Additionally, this means that relatively few sample orientations are needed in
@@
 CrystalPlan will greatly increase the scientific productivity and throughput of
-TOPAZ. More users will be able to measure more samples, faster.
+TOPAZ and other beamlines. More users will be able to measure more samples in
+less time.
 
@@
 goniometers. A base Goniometer class is subclassed for specific types of
-goniometers, allowing a unified API: in a way, it is a “universal goniometer”,
-which uses the conventional goniometer angles of $\phi$, $\chi$, $\omega$ as
-defined in \cite{knuth84}. Each class inheriting from Goniometer can convert
+goniometers, allowing a unified API. This base class acts as a \emph{universal
+goniometer},
+which uses the conventional goniometer angles of
+$\phi$, $\chi$, $\omega$ as defined in \cite{busing67}.
+Each class inheriting from Goniometer can convert
 the desired phi, chi, omega angles to the internal motor positions it requires. The API allows
@@
 
-The data analysis program ISAW is used to calculate the the sample mounting
-orientation UB matrix, which is saved to a text file along with the lattice
+From an initial data set, the data analysis program ISAW \cite{Tao2006} is used
+to calculate the the sample mounting orientation UB matrix, which is saved to a text file along with the lattice
 parameters. CrystalPlan can load this UB matrix file and will use that UB matrix
@@
 
-\subsection{Volume coverage}
-
-\subsubsection{Reciprocal Space Representation}
+\subsection{Reciprocal Space Representation.}
 
@@
 
-\subsubsection{Calculating Coverage of One Sample Orientation}
+\subsection{Calculating Coverage of One Sample Orientation}
 
@@
 measurement redundancy. A list of $n$ $3\times3$ multiplication matrices is
-generated based on the point group; any {hkl} vector can be multiplied by these matrices
-to find the n equivalent hkl reflections; for example, for orthorombic (mmm)
-symmetry, $n=8$ and 8 multiplication matrices are generated, which multiply
-{hkl} to give {±h, ±k, ±l}.
+generated based on the point group; any $hkl$ vector can be multiplied by
+these matrices to find the $n$ equivalent $hkl$ reflections; for example, for
+orthorombic (mmm) symmetry, $n=8$ and 8 multiplication matrices are generated, which multiply
+$hkl$ to give $\pm h$, $\pm k$, $\pm l$.
 
@@
 
-In volume coverage mode, a 4D matrix is generated where each voxel has a list of
-n corresponding equivalent voxels. This allows CrystalPlan to quickly compute a
-volume coverage map that takes into account crystal symmetry: at each voxel, n
-original voxels are checked to see if any of them were measured.
+In volume coverage mode, a 4D matrix is generated where each voxel has a list
+with $n$ entries, each pointing to a corresponding equivalent voxels. This
+allows CrystalPlan to quickly compute a volume coverage map that takes into
+account crystal symmetry: at each voxel, all $n$ voxels are checked to
+see if any of them were measured.
 
@@
 
-\subsubsection{Design}
+\subsubsection{Design.}
 
@@
 
-\subsubsection{Main Program Workflow}
+\subsubsection{Main Program Workflow.}
 Some screenshots of the GUI are shown in Figure \ref{fig:screenshots}. The main
@@
 
-\subsubsection{Reciprocal Space Coverage Visualization}
+\subsubsection{Reciprocal Space Coverage Visualization.}
 Figure \ref{fig:volume_view}a) shows a screenshot of the reciprocal space

### 2010-09-21 docs/CrystalPlan_paper/paper.tex
Refs #182: Added ewald sphere figure and script to generate it

@@
      \journalonline{0 XXXXXXX 0000}
+
+\usepackage[pdftex]{graphicx}
 
@@
 
-The program needs to be cross-platform, and run equally well on Linux, Mac OS
-and Windows, although Linux is our primary target operating system. Another
-design goal was to use of open-source software and toolkits as much as possible.
+Crystal plan is designed with these capabilities:
 
-The calculation code of CrystalPlan is placed in a model layer and kept
-completely separate from eventual GUI elements. This makes it possible to run
-calculations using Python scripts, if desired. Unit tests for each part of the
-calculations are used to ensure consistent results on different platforms and
-systems.
+\begin{itemize}
+  \item Load an instrument geometry.
+  \item Calculate the orientation of a crystal at different goniometer settings.
+  \item Calculate the coverage of these orientations in reciprocal space.
+  \item Take crystal symmetries into account.
+  \item Optimize the experiment plan.
+\end{itemize}
 
@@
 a text file containing the location, size, orientation and number of pixels of
-each detector.
+each detector. Other settings to be set for an instrument include the
+detectable range of neutron/x-ray wavelength.
 
@@
 goniometers. A base Goniometer class is subclassed for specific types of
-goniometers, allowing a unified API. This base class acts as a \emph{universal
-goniometer},
+goniometers, allowing a unified interface. This base class acts as a
+\emph{universal goniometer},
 which uses the conventional goniometer angles of
@@
 matrix is also specified (in $\AA^{-1}$), which determines the number of points to be
-modeled. Machine memory limits the resolution that can be achieved.
+modeled. Available machine memory limits the resolution that can be achieved in
+practice.
 
@@
 the scattered beam is known. Then, two incident beam vectors are considered,
-corresponding to the low and high wavelength limits of the detectors and/or
-source. The scattered beam direction and the two possible incident beam vectors
-are used to to find two q vectors, which point in the same direction. At the
-high wavelength we find qmin, and the low wavelength limit scattering off of qmax.
-Since the instrument has a wide bandwidth, we fill in the 3D q-space matrix by
-filling all points from qmin to the  qmax, for all pixels in the face of the
-detector. This quickly defines a volume of coverage, as shown in Figure
-\ref{fig:volume_coverage}.
+corresponding to the $\lambda_{min}$ and $\lambda_{max}$ wavelength limits
+of the detectors and/or source. The scattered beam direction and the two possible incident beam vectors
+are used to to find two $q$ vectors, which point in the same direction. At the
+high wavelength we find $q_{min}$, and the low wavelength limit scattering off
+of $q_{max}$. Since the instrument has a wide bandwidth, we fill in the 3D q-space
+matrix by filling all points from $q_{min}$ to the $q_{max}$, for all pixels in
+the face of the detector. This quickly defines a volume of coverage, as shown
+schematically in two dimensions in Figure \ref{fig:ewald_fig}.
 
@@
 $\gamma$) are typed in or loaded from the ISAW UB matrix file. CrystalPlan then
-generates a list of HKL peaks with d-spacings larger than the $d_{min}$ specified. For each
-sample orientation being simulated and for each peak, a function calculates at
-what wavelength lambda and in which direction it scatters. If the wavelength is
-within range of the limits of the instrument, the beam direction is projected
+generates a list of $hkl$ peaks with d-spacings larger than the $d_{min}$
+specified. For each sample orientation being simulated and for each peak, a function calculates at
+what wavelength $\lambda$ and in which direction it scatters. If $\lambda$ 
+is within range of the limits of the instrument, the beam direction is projected
 onto each detector face to determine where, on each detector surface, it will be
@@
 
-Each HKL peak is a separate object that holds a list of simulated measurements,
-including which detector measured it, at what position on the detector face, at
-what wavelength, and for which sample orientation. These data will be used to
-graphically display coverage and calculate statistics such as redundancy.
+Each $hkl$ peak is a separate object that holds a list of predicted
+measurements, including which detector measured it, at what position on the
+detector face, at what wavelength, and for which sample orientation. These data
+will be used to  graphically display coverage and calculate statistics such as
+redundancy.
 
@@
 
-For the single reflections, each hkl is tracked separately, but a primary hkl is
-found that points to a list of equivalent hkls. The GUI can then display either
-all hkl, ignoring crystal symmetry; or combine them to display only part of the
-reflections.
+For the single reflections, each $hkl$ is tracked separately, but a primary
+$hkl$ is found that points to a list of equivalent $hkl$'s. The GUI can then display
+either all $hkl$, ignoring crystal symmetry; or combine them to display only part of
+the reflections.
 
@@
 The initial population is created randomly and is made to evolve using mutations
-and crossover. For each individual set of sample orientations, the coverage
-statistics are computed (taking crystal symmetry into account if desired), and
-this is used as the fitness of the given individual in the normal genetic
+and crossover. Individuals whose goniometer angles turn out to be impossible to
+reach are eliminated. For each individual set of sample orientations, the
+coverage statistics are computed (taking crystal symmetry into account if desired), and this is used as the fitness of the given individual in the normal genetic
 algorithm process.
@@
 
-CrystalPlan was written in Python, a powerful, open, cross-platform scripting
-language. Python was chosen for its ease of cross-platform deployment, and for
+The program needs to be cross-platform, and run equally well on Linux, Mac OS
+and Windows, although Linux is our primary target operating system. Another
+design goal was to use of open-source software and toolkits as much as possible.
+
+To this end, CrystalPlan was written in Python, a powerful, open, cross-platform
+scripting language. Python was chosen for its ease of cross-platform deployment, and for
 its attractive language features. The numpy and scipy libraries (two
@@
 the pure Python version (which can still be used, in case of incompatibilities).
+
+\subsection{Code Layout}
+
+The calculation code of CrystalPlan is placed in a model layer and kept
+completely separate from eventual GUI elements. This makes it possible to run
+calculations using Python scripts, if desired. Unit tests for each part of the
+calculations are used to ensure consistent results on different platforms and
+systems.
 
@@
 \begin{figure}
-\caption{volume coverage}
-%\includegraphics{fig1.ps}
-\label{fig:volume_coverage}
+\caption{Ewald sphere construction showing reciprocal-space detector coverage
+for a wide-bandwidth Laue diffractometer. The smaller blue circle represents the
+$1/\lambda_{max}$ limit of measurement, with the larger green circle
+representing $1/\lambda_{min}$. The origin of reciprocal space (000) is common to both
+spheres, but the sample position in the construction (S1 or S2) depends on
+$\lambda$. The angular coverage of the detector is represented by the dashed
+wedge, and projects onto each sphere of radius $1/\lambda$ an arc. The red wedge
+between the two $\lambda$ is built by projecting straight lines between the
+$q_{min}$ and $q_{max}$ limits, giving the volume in reciprocal space that is
+measured at a single sample oriention. }
+%\includegraphics[bb=0in 0in 2in 2in, width=2in]{ewald_fig.pdf} 
+\label{fig:ewald_fig} 
 \end{figure}
 
-\
 \begin{figure}
@@
 
-\begin{figure}
-\caption{figureeeeee}
-%\includegraphics{fig1.ps}
-\label{fig:screenshots}
-\end{figure}
-
 \end{document}                    % DO NOT DELETE THIS LINE

### 2010-09-21 docs/CrystalPlan_paper/ewald_fig.py

@@ -1,0 +1,84 @@
+""" Make an Ewald sphere figure """
+#from matplotlib import rc
+#rc('font',**{'family':'sans-serif','sans-serif':['Helvetica']})
+### for Palatino and other serif fonts use:
+##rc('font',**{'family':'serif','serif':['Palatino']})
+#rc('text', usetex=True)
+
+from pylab import *
+
+clf()
+
+theta = arange(0, 2*pi, 0.01)
+
+i1 = 100
+i2 = 150
+det = range(i1,i2)
+
+cx = cos(theta)
+cy = sin(theta)
+
+lam_min = 1.0
+L1 = 1/lam_min
+lam_max = 2.5
+L2 = 1/lam_max
+
+cx1 = cx/lam_min + L1
+cy1 = cy/lam_min
+cx2 = cx/lam_max + L2
+cy2 = cy/lam_max
+
+plot( cx1, cy1, '-g')
+plot( cx2, cy2, '-b')
+axis('equal')
+
+plot( cx1[det], cy1[det], '-r', linewidth=6)
+plot( cx2[det], cy2[det], '-r', linewidth=6)
+
+plot( 0, 0, 'k.', markersize=10)
+annotate(' (000)', (0,0) , va='center')
+
+#Volume wedge
+num = i1
+plot( [cx2[num], cx1[num]], [cy2[num], cy1[num]], '-r', linewidth=2)
+num = i2
+plot( [cx2[num], cx1[num]], [cy2[num], cy1[num]], '-r', linewidth=2)
+
+annotate('$q_{min}$', xy=(cx2[num], cy2[num]), xytext=(0.3, 0.5),
+            arrowprops=dict(facecolor='black', arrowstyle='->') )
+
+annotate('$q_{max}$', xy=(cx1[num], cy1[num]), xytext=(0.8, 1.1),
+            arrowprops=dict(facecolor='black', arrowstyle='->') )
+
+r = 0.125
+plot( cx[det]*r + L2, cy[det]*r, ':k')
+r = 0.25
+plot( cx[det]*r + L1, cy[det]*r, ':k')
+annotate('angular coverage \nof detector', xy=(1.1,0.3), xytext=(1.2, 0.15),
+            arrowprops=dict(facecolor='black', arrowstyle='->') )
+
+
+#Detector wedge
+annotate('S2', (L1,0) , va='top', ha='center')
[…]

### 2010-09-21 docs/CrystalPlan_paper/ewald_fig.py
Refs #182: Had to use .ps images. Added a screenshot

@@ -8,7 +8,8 @@
 from pylab import *
 
 clf()
-
+# Fill up the axes
+axes([0,0,1,1])
 theta = arange(0, 2*pi, 0.01)
 
 i1 = 100
@@ -81,4 +82,5 @@ yticks( [])
 
 savefig("ewald_fig.png", dpi=300)
 savefig("ewald_fig.pdf")
+savefig("ewald_fig.ps")
 

### 2010-09-21 docs/CrystalPlan_paper/paper.tex
Refs #182: Had to use .ps images. Added a screenshot
Refs #182: Extra screenshot.

@@
 
-\documentclass{iucr}              % DO NOT DELETE THIS LINE
+\documentclass[final]{iucr}              % DO NOT DELETE THIS LINE
 
@@
 
-\usepackage[pdftex]{graphicx}
+\usepackage{graphicx}
+%\usepackage[pdftex]{graphicx}
+%\DeclareGraphicsExtensions{.pdf,.png,.jpg}
 
@@
 \subsubsection{Main Program Workflow.}
-Some screenshots of the GUI are shown in Figure \ref{fig:screenshots}. The main
-window consists of a series of tabs, with workflow proceeding from left to right:
+A screenshot of the main window GUI is shown in Figure \ref{fig:screenshots}.
+The main window consists of a series of tabs, with workflow proceeding
+intuitively from left to right:
 
@@
   	\item Q-Space tab: define the volume of interest (minimum d-spacing) and
-resolution of reciprocal space. Detectors tab: load instrument detector geometry
-files, and enable or disable individual detectors. Goniometer tab: to choose the
+resolution of reciprocal space. 
+ 	\item Detectors tab: load instrument detector geometry
+files, and enable or disable individual detectors. 
+	\item Goniometer tab: to choose the
 goniometer and degrees of orientation freedom of the instrument.
-
-  \item Sample tab: enter or load the crystal's lattice parameters and UB matrix.
-
+  	\item Sample tab: enter or load the crystal's lattice parameters and UB
+  matrix.
   	\item Try an Orientation tab: interactively rotate a sample and observe the
   changes in coverage in real-time.
+  	\item Add Orientations tab: type in a list of sample orientation angles to
+  add to the experiment plan.
 
-  \item Add Orientations tab: type in a list of sample orientation angles to add
-  to the experiment plan.
-
-  \item Experiment Plan tab: a list of the previously calculated sample orientations.
-Here you can manage the orientations you will use in the experiment: add,
+  	\item Experiment Plan tab: a list of the previously calculated sample
+orientations. Here you can manage the orientations you will use in the experiment: add,
 delete, or temporarily disable them to see the effect on coverage. Once
@@
 coverage 3D interface. The 3D view can represent q-space in two ways: volume coverage, or single-peak
-coverage.
+coverage. The view shown can be rotated, translated or zoomed using the mouse or
+keyboard.
 
@@
 
-The reflections view (Figure \ref{fig:refl_view}a) displays each hkl as a
-small sphere; the color of which indicates the number of times the hkl is predicted to be measured
-in the current experiment plan.
+The reflections view (Figure \ref{fig:refl_view}a) displays each $hkl$ as a
+small sphere; the color of which indicates the number of times the $hkl$ is
+predicted to be measured in the current experiment plan.
 
-Both views can take into crystal symmetry in their display (Figures
-\ref{fig:volume_view}b and \ref{fig:refl_view}b), using the techniques described
-previously.
+Both views can take into crystal symmetry in their display using the techniques
+described previously. 
+The volume coverage (Figures \ref{fig:volume_view}b) fills in the sphere of
+coverage; 
+whereas the single reflection view (Figure \ref{fig:refl_view}b) leaves only
+the primary reflections (one eight of the total, in this example).
 
-Finally, both views display quick coverage statistics, showing the percentage of
-volume (or of the number of peaks) measured, as well as the proportion of
-redundancy (voxels or peaks measured 2 or more times).
+Finally, both views display quick coverage statistics in the lower right
+corner, showing the percentage of volume (or of the number of peaks) measured,
+as well as the proportion of redundancy (voxels or peaks measured 2 or more
+times).
+
+\subsubsection{Single Reflection View.}
+Figure \ref{fig:single_reflection} shows a screenshot of the GUI allowing one to
+observe in detail the predicted measurements of a particular peak, by clicking
+in the 3D view or by typing in the $hkl$ values. The position on the detector
+and wavelength of measurement are displayed. An additional window (not shown) allows the user to place the peak on a particular
+position on a detector by automatically finding the necessary sample orientation
+angles. This is used, for example, to center important peaks in the most
+sensitive part of a detector.
 
@@
 
-Live data processing    …  interface with ISAW … InelasCasetic instruments ??
+We have described CrystalPlan, a software program designed to quickly produce
+efficient experiment plans for wide bandwidth time-of-flight Laue
+diffractometers. This program will help maximize scientific productivity at SNS
+beamlines such as TOPAZ.
+
+One planned future development for CrystalPlan include extending the software
+to help predict coverage in inelastic neutron scattering experiments. Another
+desired feature is the introduction of feedback loops to help determine the
+right time to end data acquisition at a particular sample orientation.
+CrystalPlan would interface with data analysis software such as ISAW and use
+recently developped capabilities for real-time data analysis to monitor
+the  signal to noise ratio on particular peaks. When the signal becomes
+sufficient, the sample will automatically be moved to the next sample
+orientation, further minimizing data acquisition time. Combined with advanced
+sample changers coming on-line at TOPAZ, this will allow for extremely
+efficient, unattended data acquisition.
 
@@
 measured at a single sample oriention. }
-%\includegraphics[bb=0in 0in 2in 2in, width=2in]{ewald_fig.pdf} 
+
+%\includegraphics[bb=0in 0in 8in 6in, scale=0.20]{ewald_fig.pdf}
+\includegraphics[width=3.5in]{ewald_fig.ps}
+
 \label{fig:ewald_fig} 
@@
 \begin{figure}
-\caption{Screenshots of the main GUI}
-%\includegraphics{fig1.ps}
+\caption{Screenshot of the main GUI window. Six more interface windows (not
+shown) are presented to the user when he or she clicks on the tabs at the
+top of the main window.}
+\includegraphics{main_gui.eps}
 \label{fig:screenshots}
@@
 
+\begin{figure}
+\caption{GUI showing details of the predicted measurement positions for a
+single $hkl$.}
+\includegraphics[width=2.5in]{single_reflection_2.eps}
+\label{fig:single_reflection}
+\end{figure}
+
 \end{document}                    % DO NOT DELETE THIS LINE

### 2010-09-22 docs/CrystalPlan_paper/paper.tex
Refs #182: RMS stats for peak offsets, more paper writing.

@@
 
-\documentclass[final]{iucr}              % DO NOT DELETE THIS LINE
+\documentclass[draft]{iucr}              % DO NOT DELETE THIS LINE
 
@@
 \usepackage{graphicx}
+\usepackage{gensymb}
+\usepackage[version=3]{mhchem} 
 %\usepackage[pdftex]{graphicx}
@@
 orientation (the U matrix loaded from ISAW) and the goniometer rotation
-($\Phi$, $\Xi$, $\Omega$) is computed. The coverage in reciprocal space is
+($\Phi$, $\chi$, $\Omega$) is computed. The coverage in reciprocal space is
 computed in this way: for each pixel in the face of a detector, the direction of
@@
 
+%---------------------------------------------------------------------------------------
+%---------------------------------------------------------------------------------------
+%---------------------------------------------------------------------------------------
+
 \section{Sample Use Case}
+
+As part of the TOPAZ beamline commissioning process, a Natrolite
+(\ce{Na2Al2Si3O10} $\cdot$ \ce{2H2O}) single crystal was measured on a fixed
+$45\degree\ \chi$ goniometer at room temperature and ambient atmosphere. An initial short run was processed using
+ISAW to determine the UB matrix and cell parameters of the crystal.
+
+The resulting UB matrix was loaded into CrystalPlan, where the Automatic
+Coverage Optimizer was used to generate an experiment plan with 8 sample
+orientations that covered 95\% of the peaks, taking into account the crystal
+symmetry (orthorombic or mmm). The plan was left to run unattended overnight.
+
+Peaks were integrated using ISAW, and the resulting .integrate file was loaded
+into CrystalPlan for comparison. The predicted and actual measurement
+positions on each detector face were compared for the 554 integrated peaks
+found. Very good agreement was found; the
+RMS difference between the predicted and actual measurement wavelength was only
+0.016 $\AA$, whereas the RMS of the detector position error was
+7.6 mm. Inspecting the offsets showed that the main contributor to this error was a
+slight misalignment of the goniometer, which CrystalPlan can help diagnose.
+
+%NOTE: Check on this misalignment of goniometer business - is that really the
+% cause?
+
+%---------------------------------------------------------------------------------------
+%---------------------------------------------------------------------------------------
+%---------------------------------------------------------------------------------------
 
@@
 
-\begin{table}
-\caption{Caption to table}
-\begin{tabular}{llcr}      % Alignment for each cell: l=left, c=center, r=right
- HEADING    & FOR        & EACH       & COLUMN     \\
-\hline
- entry      & entry      & entry      & entry      \\
- entry      & entry      & entry      & entry      \\
- entry      & entry      & entry      & entry      \\
-\end{tabular}
-\end{table}
-
  % Postscript figures can be included with multiple figure blocks
@@
 \begin{figure}
-\caption{Ewald sphere construction showing reciprocal-space detector coverage
+\caption{2D Ewald sphere construction showing reciprocal-space detector coverage
 for a wide-bandwidth Laue diffractometer. The smaller blue circle represents the
@@
 representing $1/\lambda_{min}$. The origin of reciprocal space (000) is common to both
-spheres, but the sample position in the construction (S1 or S2) depends on
-$\lambda$. The angular coverage of the detector is represented by the dashed
-wedge, and projects onto each sphere of radius $1/\lambda$ an arc. The red wedge
-between the two $\lambda$ is built by projecting straight lines between the
-$q_{min}$ and $q_{max}$ limits, giving the volume in reciprocal space that is
-measured at a single sample oriention. }
+spheres, but the ``virtual" sample position in the construction (S1 or S2)
+depends on $\lambda$. The angular coverage of the detector is represented by the dashed
+wedge, and projects an arc onto each sphere of radius $1/\lambda$. The
+red wedge between the two $\lambda$ is built by projecting straight lines between the
+$q_{min}$ and $q_{max}$ limits for each pixel on the detector face, giving the
+volume in reciprocal space that is measured at a single sample oriention. }
 

### 2010-09-22 model/tools.py

@@ -21,12 +21,13 @@ from numpy_utils import column, rotation_matrix, vector_length, normalize_vector
 #================================================================================
 class PeakOffset:
     """Class holds info about the offset between predicted and real measurement position"""
-    def __init__(self, det_num, wavelength, pred_h, pred_v, real_h, real_v):
+    def __init__(self, det_num, wavelength_predicted, wavelength_measured, pred_h, pred_v, real_h, real_v):
         self.det_num = det_num
         self.predicted = np.array([pred_h, pred_v])
         self.measured = np.array([real_h, real_v])
         self.offset = self.measured-self.predicted
-        self.wavelength = wavelength
+        self.wavelength_predicted = wavelength_predicted
+        self.wavelength_measured = wavelength_measured
         
     def __str__(self):
         return "Det %d offset of %.3f, %.3f; wl=%.3f." % (self.det_num, self.offset[0], self.offset[1], self.wavelength)
@@ -61,7 +62,7 @@ def calculate_peak_offsets():
                             pred_angles = np.array(gon.get_phi_chi_omega(poscov.angles))
                             if np.allclose(pred_angles, np.array(rrm.angles), atol=1e-2):
                                 #Great, you found a match
-                                po = PeakOffset(rm.detector_num, rrm.wavelength, rm.horizontal, rm.vertical, rrm.horizontal, rrm.vertical)
+                                po = PeakOffset(rm.detector_num, rm.wavelength, rrm.wavelength, rm.horizontal, rm.vertical, rrm.horizontal, rrm.vertical)
                                 offsets.append(po)
                                 #print po
 
@@ -76,6 +77,23 @@ def plot_peak_offsets(offsets, filebase, doshow=False):
     #@type inst Instrument
     inst = instrument.inst
     numperpage = 6
+
+    # Initialize some stats
+    rms = 0
+    rms_wl = 0
+    #@type po PeakOffset
+    for po in offsets:
+        #Square of the error distance
+        error = (po.measured[0]- po.predicted[0])**2 + (po.measured[1]- po.predicted[1])**2
+        rms += error
+        error = (po.wavelength_measured - po.wavelength_predicted)**2
+        rms_wl += error
+    # Now do the root-mean
+    rms = (rms/ len(offsets))**0.5
[…]

### 2010-09-22 docs/CrystalPlan_paper/paper.tex
Refs #182: version of the paper I will send around for comments.

@@
 
-\documentclass[draft]{iucr}              % DO NOT DELETE THIS LINE
+\documentclass[final]{iucr}              % DO NOT DELETE THIS LINE
 
@@
 \usepackage{graphicx}
-\usepackage{gensymb}
-\usepackage[version=3]{mhchem} 
+\usepackage{gensymb} %General symbols
+\usepackage[version=3]{mhchem} %For chemical formulae
+\usepackage{subfigure}
+
 %\usepackage[pdftex]{graphicx}
 %\DeclareGraphicsExtensions{.pdf,.png,.jpg}
+
+%quick angstrom define
+\newcommand{\ang}{$\mathrm{\AA} $}
 
@@
 
-Crystal plan is designed with these capabilities:
+CrystalPlan is designed with these capabilities:
 
@@
   \item Calculate the orientation of a crystal at different goniometer settings.
-  \item Calculate the coverage of these orientations in reciprocal space.
+  \item Calculate the coverage of these orientations in reciprocal space,
+  in volume and for single peaks.
   \item Take crystal symmetries into account.
-  \item Optimize the experiment plan.
+  \item Optimize the experiment plan automatically.
 \end{itemize}
@@
 of which is determined by the user, who specifies the minimum d-spacing
-($d_{min}$, in $\AA$) that he or she is interested in. The resolution of the 3D
-matrix is also specified (in $\AA^{-1}$), which determines the number of points to be
-modeled. Available machine memory limits the resolution that can be achieved in
+($d_{min}$, in \ang) that he or she is interested in. The resolution of the
+3D matrix is also specified (in \ang$^{-1}$), which determines the number of
+points to be modeled. Available machine memory limits the resolution that can be achieved in
 practice.
@@
 these matrices to find the $n$ equivalent $hkl$ reflections; for example, for
-orthorombic (mmm) symmetry, $n=8$ and 8 multiplication matrices are generated, which multiply
-$hkl$ to give $\pm h$, $\pm k$, $\pm l$.
+orthorombic (mmm) symmetry, $n=8$; therefore 8 multiplication matrices are
+generated that multiply $hkl$ to give $\pm h$, $\pm k$, $\pm l$.
 
@@
 (\ce{Na2Al2Si3O10} $\cdot$ \ce{2H2O}) single crystal was measured on a fixed
-$45\degree\ \chi$ goniometer at room temperature and ambient atmosphere. An initial short run was processed using
-ISAW to determine the UB matrix and cell parameters of the crystal.
+$45\degree\ \chi$ goniometer at room temperature and ambient atmosphere. An
+initial short run (wavelength range from $\sim0.5$ \ang\ to $3.5$ \ang)
+was processed using ISAW to determine the UB matrix and cell parameters of the crystal.
 
@@
 RMS difference between the predicted and actual measurement wavelength was only
-0.016 $\AA$, whereas the RMS of the detector position error was
+0.016 \ang, whereas the RMS of the detector position error was
 7.6 mm. Inspecting the offsets showed that the main contributor to this error was a
@@
 \begin{figure}
-\caption{Volume coverage in reciprocal space view: a) without crystal symmetry;
-b) with crystal symmetry}
-%\includegraphics{fig1.ps}
+\caption{Sample screenshot showing the 3D view of the predicted coverage of
+reciprocal space volume for 3 sample orientations and 14 detectors in TOPAZ, from 0.9 \ang\ to 4.0\ang:
+a) without crystal symmetry, 17\% coverage; b) with crystal symmetry (mmm 8-fold
+symmetry), coverage increases to 65\%. }
+\includegraphics{volume_nosym.eps}
+a)
+\includegraphics{volume_sym.eps}
+b)
 \label{fig:volume_view}
@@
 \begin{figure}
-\caption{Coverage of single-crystal reflections in reciprocal space: a) without
-crystal symmetry; b) with crystal symmetry}
-%\includegraphics{fig1.ps}
+\caption{Sample screenshot showing the 3D view of single-crystal reflection
+measurements. Each sphere represents an $hkl$ peak, color-coded to indicate the
+number of measurements (dark blue: 0, cyan: 1, green: 2, etc.). 
+a) without crystal symmetry, 2375 of 9481 reflections are measured, or 25\%.
+b) with crystal symmetry, 1276 of 1427 primary reflections are measured, or
+89\%, and redundancy is also increased. }
+\includegraphics{refl_nosym.eps}
+a)
+\includegraphics{refl_sym.eps}
+b)
 \label{fig:refl_view}

### 2010-09-24 docs/CrystalPlan_paper/paper.tex
Refs #182: Xiaoping's revisions and new figures for CrystalPlan paper.

@@
 productivity. 
-An attractive GUI including a 3D viewer and an automated coverage optimizer 
+A user friendly GUI including a 3D viewer, an automated coverage optimizer,
+and an option to reorient the crystal for the measurement 
+of selected hkls on specific detector positions 
 are among its useful features. 
@@
 Lab (ORNL), currently finishing commissioning, is a neutron time-of-flight
-instrument designed to acquire single-crystal diffraction data of
-small-to-moderate sized unit cells with high throughput. The final design calls
-for an array of 48 detectors covering a large fraction of  $2\pi$ steradians;
-however as of this writing, only 14 of the 48 detectors are installed. In order
-to have sufficient data to perform a fit, an experimenter will  typically want
+Laue instrument with high wavelength resolution. It is designed to acquire
+single-crystal diffraction data of small-to-moderate sized unit cells with  high
+throughput. The final design calls for an array of 48 detectors covering a large
+fraction of $2\pi$ steradians. However as of this writing, only 14 of the 48
+detectors are installed, giving only a fraction of $2\pi$ steradian coverage. 
+In order to have sufficient data to perform a fit, an experimenter will typically want
 to measure $ > 85\%$ of the peaks within a certain range of d-spacings. 
-However, TOPAZ's complex geometry makes it very difficult to come up with an
-effective set of sample orientations that will cover all the required peaks
-without unnecessary redundancy.
+However, wide wavelength band combined with TOPAZ's complex
+geometry makes it very difficult to come up with an effective set of sample
+orientations that will cover all the required peaks without unnecessary redundancy.
 
 Existing diffraction experiment planning software programs (for example,
-CrysAlis Pro from Oxford Diffraction; DIFFRAC from Bruker AXS) were designed for
+CrysAlis Pro from Oxford Diffraction; COSMO from Bruker AXS) were designed for
 X-ray experiments where a single  incident wavelength and a large number of
 sample orientations or sweeps are used to collect data. 
-These programs cannot be used for TOPAZ, since as a time-of-flight instrument, 
-it is able to collect data from a wide bandwidth of incident neutrons. 
+These programs are not suitable for TOPAZ, since as a time-of-flight instrument, 
+it is able to collect data from a wide bandwidth of
+incident neutrons, using a stationary crystal. 
 Additionally, this means that relatively few sample orientations are needed in
@@
 Coverage Optimizer was used to generate an experiment plan with 8 sample
-orientations that covered 95\% of the peaks, taking into account the crystal
-symmetry (orthorombic or mmm). The plan was left to run unattended overnight.
+orientations (Figure \ref{fig:natrolite_gui}a) that covered 97\% of the peaks,
+taking into account the crystal symmetry (orthorombic or mmm) (Figure
+\ref{fig:natrolite_gui}b). The plan was left to run unattended overnight.
 
@@
 RMS difference between the predicted and actual measurement wavelength was only
-0.016 \ang, whereas the RMS of the detector position error was
-7.6 mm. Inspecting the offsets showed that the main contributor to this error was a
-slight misalignment of the goniometer, which CrystalPlan can help diagnose.
+0.016 \ang. Inspecting the detector position offsets showed that the main
+contributor to this error was a slight misalignment of the goniometer, which
+CrystalPlan can help diagnose.
+
+%, whereas the RMS of the detector position error was 7.6 mm.
 
@@
 
+\begin{figure}
+\caption{a) Screenshot of the Natrolite experiment plan generated using the
+Automatic coverage Optimizer. b) Screenshot of the resulting coverage
+statistics.}
+\includegraphics[width=3.5in]{Plan8List.eps}
+a)
+\includegraphics[width=3.5in]{NatCoverage.eps}
+b)
+\label{fig:natrolite_gui}
+\end{figure}
+
 \end{document}                    % DO NOT DELETE THIS LINE

### 2010-11-04 gui/dialog_goniometer_angles.py
Fixes #186: Improved dialog for advanced goniometer angle editing.

@@ -1,0 +1,216 @@
+"""
+GUI to set crystal parameters, as a dialog.
+"""
+# Author: Janik Zikovsky, zikovskyjl@ornl.gov
+# Version: $Id: dialog_edit_crystal.py 1424 2010-10-18 17:38:30Z 8oz $
+
+#--- General Imports ---
+import wx
+import os
+import copy
+import numpy as np
+import sys
+
+#--- Traits imports ---
+from enthought.traits.api import HasTraits,Int,Float,Str,String,Property,Bool, List, Tuple, Array
+from enthought.traits.ui.api import View,Item,Group,Label,Heading, Spring, Handler, TupleEditor, TabularEditor, ArrayEditor, TextEditor, CodeEditor
+from enthought.traits.ui.editors import EnumEditor
+from enthought.traits.ui.menu import OKButton, CancelButton
+
+#--- GUI Imports ---
+
+#--- Model Imports ---
+if __name__=="__main__":    sys.path.insert(0, "..")
+import model
+from model.crystals import Crystal
+
+# ===========================================================================================
+class GoniometerEditorTraitsHandler(Handler):
+    """Handler that reacts to changes in the goniometer object."""
+
+    def __init__(self, frame, *args, **kwargs):
+        Handler.__init__(self, *args, **kwargs)
+        self.add_trait('frame', frame)
+        self.add_trait('user_clicked_okay', False)
+
+    #---------------------------------------------------------------------------
+    def setattr(self, info, object, name, value):
+        """Called when any attribute is set."""
+        #The parent class actually sets the value
+        Handler.setattr(self, info, object, name, value)
+        #Do other checks
+        self.check_validity()
+
+    #---------------------------------------------------------------------------
+    def check_validity(self):
+        # Do nothing for now - always valid
+        pass
+
+    #---------------------------------------------------------------------------
+    def close(self, apply_change):
+        """Close """
+        if apply_change:
+            #Modify the crystal data in the experiment.
+            self.frame.original_gonio.copy_traits(self.frame.gonio)
+        #Close it
+        self.user_clicked_okay = apply_change
+        self.frame.Close()
+
+
+
+# ===========================================================================================
[…]

### 2010-11-04 gui/panel_detectors.py

@@ -15,10 +15,10 @@ import numpy as np
 #--- GUI Imports ---
 import display_thread
 import gui_utils
+from enthought.traits.api import HasTraits, Range, Instance, on_trait_change
+from enthought.traits.ui.api import View, Item, HGroup
 try:
     from mlab_utils import *
-    from enthought.traits.api import HasTraits, Range, Instance, on_trait_change
-    from enthought.traits.ui.api import View, Item, HGroup
     from enthought.tvtk.pyface.scene_editor import SceneEditor
     from enthought.mayavi.tools.mlab_scene_model import MlabSceneModel
     from enthought.mayavi.core.ui.mayavi_scene import MayaviScene

### 2010-11-04 gui/panel_goniometer.py

@@ -10,6 +10,7 @@ goniometer.
 #--- General Imports ---
 import wx
 import copy
+import dialog_goniometer_angles
 
 #--- GUI Imports ---
 import gui_utils
@@ -41,17 +42,15 @@ class PanelGoniometerController():
         #Select the 1st one by default
         self.selected = model.goniometer.goniometers[0]
 
-
-    #------------------------------------------------------------------
-    def update_current(self):
-        gon = copy.copy(model.instrument.inst.goniometer)
-        self.current_gon_copy = gon
+    def show_current_gon_copy(self):
+        """ Updates the GUI with whatever is in
+        self.current_gon_copy """
         if not self.panel.currentControl is None:
-            #Remove the existing one
+            # Remove the existing one
             self.panel.boxSizerAll.Remove(self.panel.currentControl)
             self.panel.currentControl.Destroy()
 
-        if gon is None:
+        if self.current_gon_copy is None:
             self.panel.staticTextCurrentGonio.SetLabel("None selected!")
         else:
             self.panel.staticTextCurrentGonio.SetLabel('') #(gon.name)
@@ -59,6 +58,13 @@ class PanelGoniometerController():
             self.panel.boxSizerAll.Insert(2, self.panel.currentControl, 0, flag=wx.EXPAND | wx.SHRINK)
 
         self.panel.boxSizerAll.Layout()
+
+    #------------------------------------------------------------------
+    def update_current(self):
+        # Start by making a full copy of the current goniometer
+        gon = copy.deepcopy(model.instrument.inst.goniometer)
+        self.current_gon_copy = gon
+        self.show_current_gon_copy()
 
     #------------------------------------------------------------------
     def update_selection(self):
@@ -98,14 +104,9 @@ class PanelGoniometerController():
     #------------------------------------------------------------------
     def edit_angles(self, event):
         """Open dialog to change the angles"""
-        view = View(
-            Label('In this advanced view, you can change the settings of each angle controlled by the goniometer.'),
-            Item('gonio_angles', editor=ListEditor(rows=3)),
-            Label('This "angle" controls the center of the wavelength bandwidth.'),
-            Item('wl_angles', editor=ListEditor(rows=3)),
-            buttons=[OKButton, CancelButton]
-            )
-        result = self.current_gon_copy.configure_traits(self.panel, view=view)
[…]

### 2010-11-05 model/goniometer.py
Fixes #180: Incorrect detectorPlacer if you change the chi angle on your goniometer

@@ -540,10 +540,12 @@ class LimitedGoniometer(Goniometer):
     def __init__(self, wavelength_control=False, *args, **kwargs):
         Goniometer.__init__(self, wavelength_control, *args, **kwargs)
 
+    #-------------------------------------------------------------------------
+    def get_fitness_function_c_code(self):
         #C code for the fitness of phi,chi, omega.
         # OVERWRITE THIS FOR SUBCLASSES!
         #   Don't change the # of parameters - the search code always gives you phi, chi, omega.
-        self.fitness_function_c_code = """
+        return """
         FLOAT fitness_function(FLOAT phi, FLOAT chi, FLOAT omega)
         {
             return absolute(chi) + absolute(omega) + absolute(phi)/10000.0;
@@ -660,8 +662,9 @@ class LimitedGoniometer(Goniometer):
             else { return value; }
         }
         """
+        
         #Add the function for the fitness
-        support += self.fitness_function_c_code
+        support += self.get_fitness_function_c_code()
 
         code = """
         FLOAT rot_angle;
@@ -774,7 +777,7 @@ class LimitedGoniometer(Goniometer):
         """
         #Workaround for bug in weave, where it ignores any changes in the support code.
         code += "\n\n // " + self.__class__.__name__ + "\n"
-        code += "/* " + self.fitness_function_c_code + " */"
+        code += "/* " + self.get_fitness_function_c_code() + " */"
 
         #List of fitnesses
         fitnesses = []
@@ -924,6 +927,18 @@ class TopazAmbientGoniometer(LimitedGoniometer):
             (self.chi == other.chi)
 
     #-------------------------------------------------------------------------
+    def get_fitness_function_c_code(self):
+        """Generate a bit of C code that gives the fitness of phi,chi, omega"""
+        s = """
+        FLOAT fitness_function(FLOAT phi, FLOAT chi, FLOAT omega)
+        {
+            // #Chi needs to be exactly this many radians
+            return absolute(chi - (%f) );
+        }
+        """ % (self.chi)
+        return s
+
+    #-------------------------------------------------------------------------
     def __init__(self, wavelength_control=False):
         """Constructor"""
         #Init the base class
@@ -941,17 +956,7 @@ class TopazAmbientGoniometer(LimitedGoniometer):
             AngleInfo('Phi'),
             AngleInfo('Omega'),
             ]
-
-        #C code for the fitness of phi,chi, omega
[…]

### 2010-11-05 gui/detector_plot.py
Refs #184: Interface to pick a specific direction for pointing an HKL

@@ -59,6 +59,7 @@ class DetectorPlot(wx.Window):
     def set_measurement(self, meas):
         """Set a measurement to be plotted"""
         self.meas = meas
+        #Extract the detector from the measurement object
         if (meas.detector_num >= 0) and (meas.detector_num < len(model.instrument.inst.detectors)):
             self.set_detector(model.instrument.inst.detectors[meas.detector_num])
         self.Refresh()
@@ -67,6 +68,12 @@ class DetectorPlot(wx.Window):
     def set_detector(self, detector):
         """Set the detector to be plotted."""
         self.detector = detector
+
+        if self.detector is None:
+            (self.det_width, self.det_height) = (150., 150.)
+        else:
+            (self.det_width, self.det_height) = (self.detector.width, self.detector.height)
+
         self.Refresh()
 
     #---------------------------------------------------------------------------
@@ -209,4 +216,8 @@ if __name__ == "__main__":
     import gui_utils
     (app, pnl) = gui_utils.test_my_gui(DetectorPlot)
     app.frame.SetClientSize(wx.Size(500, 700))
+    det = model.detectors.FlatDetector("Modified det")
+    det.width = 250
+    det.height = 125
+    pnl.set_detector(det)
     app.MainLoop()

### 2010-11-05 gui/gui_utils.py

@@ -325,7 +325,9 @@ def scale_to_fit(source, target):
         wx.Size() containing the scaled size
         ratio: what to multiply the source to get to the resized size
     """
-    ratio = max( [source.width / float(target.width), source.height / float(target.height)] )
+    ratio = max( [float(source.width) / float(target.width), float(source.height) / float(target.height)] )
+    # Avoid divide by zero
+    if (ratio == 0.0): ratio = 1.0
     return (wx.Size(source.width / ratio, source.height / ratio), 1/ratio)
 
 #---------------------------------------------------------------------------

### 2010-11-05 gui/reflection_placer.py

@@ -45,8 +45,12 @@ class ReflectionPlacerHandler(Handler):
         #The parent class actually sets the value
         Handler.setattr(self, info, object, name, value)
         self.changed_point(object)
-        #Choosing another reflection, or a different detector
-        if name == "hkl" or name=="detector" or name=="brute_search":
+        #Choosing another reflection, or a different detector,
+        # or switching to arbitrary mode
+        if name in ["hkl", "detector", "brute_search", 
+                    "arbitrary_bool", "arbitrary_xyz", "arbitrary_width"]:
+            #Make sure the plot detector points to the right one
+            self.frame.detectorPlot.set_detector(self.frame.placer.get_detector())
             #Make the map recalculate
             self.frame.map_thread.reset()
 
@@ -59,6 +63,7 @@ class ReflectionPlacerHandler(Handler):
         meas = ReflectionMeasurement(None, None)
         meas.horizontal = xy[0]
         meas.vertical = xy[1]
+        meas.detector_num = -1 # Mark to say NO detector
         self.frame.detectorPlot.set_measurement(meas)
         #Update other GUI
         valid_angle = not np.any(np.isnan(object.angles_deg))
@@ -143,7 +148,9 @@ class PlacerMapThread(Thread):
                 try:
                     #Calculate more!
                     t1 = time.time()
-                    self.frame.placer.calculate_allowed_map(xpixels, ypixels, step, allowed, wavelength_map, calculated, block_filled, self.frame.placer.brute_search, callback=self.frame.calculation_callback)
+                    self.frame.placer.calculate_allowed_map(xpixels, ypixels, step,
+                        allowed, wavelength_map, calculated, block_filled,
+                        self.frame.placer.brute_search, callback=self.frame.calculation_callback)
                     #print "step ", step, "took ", time.time()-t1, " seconds."
                     if not self._want_abort:
                         #To avoid bug when closing
@@ -199,6 +206,10 @@ class ReflectionPlacer(HasTraits):
     wavelength = Float(0.0)
     wavelength_can_be_measured = Bool(True)
 
+    arbitrary_bool =  Bool(False, desc='Arbitrary direction instead of a detector?')
+    arbitrary_xyz =  Array( shape=(1,3), dtype=float, desc='XYZ coordinates (in mm) to use as the direction to use. +Y = up; +Z = beam direction.')
[…]

### 2010-11-05 model/goniometer.py

@@ -844,7 +844,10 @@ class LimitedGoniometer(Goniometer):
         initial_R = numpy_utils.rotation_matrix_around_vector(rotation_axis, angle)
 
         result_vec = np.dot(initial_R, column(starting_vec)).flatten()
-        assert np.allclose( result_vec, ending_vec), "initial rotation matrix makes the correct rotation. Got %s, expected %s" % ( result_vec, ending_vec)
+        #Check that the matrices match, but not if all are NaN
+        #if not np.any(np.isnan(result_vec) and np.isnan(ending_vec)):
+        if not np.any(np.isnan(result_vec)):
+            assert np.allclose( result_vec, ending_vec), "initial rotation matrix makes the correct rotation. Got %s, expected %s" % ( result_vec, ending_vec)
 
 
         def optimize(start, stop, step):

### 2010-11-05 gui/panel_sample.py
Refs #185, a b c vectors displayed in the Sample tab.

@@ -191,6 +191,7 @@ class PanelSample(wx.Panel):
     def __init__(self, parent):
         self._init_ctrls(parent)
 
+        vector_format = "%.3f"
         #Make a simple, mostly read-only view for the crystal
         self.crystal_view = View(
             Item("name", label="Crystal Name"),
@@ -199,9 +200,12 @@ class PanelSample(wx.Panel):
             Item("lattice_angles_deg_arr", label="Lattice angles (degrees)", format_str="%.3f", style='readonly'),
             Item("ub_matrix", label="Sample's UB Matrix", style='readonly', format_str="%9.5f"),
             Item("point_group_name", label="Point Group", style='readonly'),
-            Item("recip_a", label="a*", style='readonly'),
-            Item("recip_b", label="b*", style='readonly'),
-            Item("recip_c", label="c*", style='readonly'),
+#            Item("recip_a", label="a*", style='readonly'),
+#            Item("recip_b", label="b*", style='readonly'),
+#            Item("recip_c", label="c*", style='readonly'),
+            Item("a", label="a vector", format_str=vector_format, style='readonly'),
+            Item("b", label="b vector", format_str=vector_format, style='readonly'),
+            Item("c", label="c vector", format_str=vector_format, style='readonly'),
             resizable=True
             )
 

### 2010-11-05 model/crystals.py

@@ -30,6 +30,7 @@ class Crystal(HasTraits):
     name = Str("name")
     description = Str("description")
 
+    # For manually specifying the lattice direction lengths
     lattice_lengths_arr = Array( shape=(1,3), dtype=np.float)
     lattice_angles_deg_arr = Array( shape=(1,3), dtype=np.float)
     #Can the lattice exist given these values?
@@ -57,6 +58,12 @@ class Crystal(HasTraits):
     recip_a = Array( shape=(3,), dtype=np.float)
     recip_b = Array( shape=(3,), dtype=np.float)
     recip_c = Array( shape=(3,), dtype=np.float)
+
+    # Real crystal lattice vectors, in XYZ coordinates.
+    a = Array( shape=(1,3), dtype=np.float)
+    b = Array( shape=(1,3), dtype=np.float)
+    c = Array( shape=(1,3), dtype=np.float)
+
     #Same info, as a column-wise a,b,c matrix
     reciprocal_lattice = Array( shape=(3,3), dtype=np.float)
 
@@ -130,6 +137,7 @@ class Crystal(HasTraits):
         the lattice volume."""
         #Check the lattice parameters.
         (a,b,c, V) = crystal_calc.make_lattice_vectors(self.lattice_lengths, self.lattice_angles)
+
         #Bad angles make a nan volume
         #   0 or negative volume also is bad
         return not ( np.isnan(V) or (V <= 1e-5) )
@@ -150,6 +158,9 @@ class Crystal(HasTraits):
         #Now the UB matrix
         self.ub_matrix = crystal_calc.make_UB_matrix(self.lattice_lengths, self.lattice_angles, phi, chi, omega)
         self.ub_matrix_is_from = "\nManually generated.\n"
+
+        # and re-calc the real-space a,b,c vectors
+        self.calculate_abc()
 
 
     #--------------------------------------------------------------------
@@ -214,6 +225,9 @@ class Crystal(HasTraits):
             angles_deg = np.rad2deg(angles)
             self.ub_matrix_is_from = "ISAW UB matrix file at\n " + filename + "\nmodified by phi, chi, omega of %.1f, %.1f, %.1f" % (angles_deg[0],angles_deg[1],angles_deg[2])
 
+            # and re-calc the real-space a,b,c vectors
+            self.calculate_abc()
+
 
     #--------------------------------------------------------------------
     def calculate_reciprocal(self):
@@ -223,6 +237,19 @@ class Crystal(HasTraits):
[…]

### 2010-12-16 model/goniometer.py
Refs #188 intermediate step

@@ -563,11 +563,23 @@ class LimitedGoniometer(Goniometer):
         #C code for the fitness of phi,chi, omega.
         # OVERWRITE THIS FOR SUBCLASSES!
         #   Don't change the # of parameters - the search code always gives you phi, chi, omega.
+        #   The *_min and *_max are the limits to these angles, in the same units.
         # This sample fitness value makes phi less important (since it normally has full freedom)
         return """
-        FLOAT fitness_function(FLOAT phi, FLOAT chi, FLOAT omega)
-        {
-            return absolute(chi) + absolute(omega) + absolute(phi)/10000.0;
+        FLOAT fitness_function(FLOAT phi, FLOAT chi, FLOAT omega,
+                FLOAT phi_min, FLOAT phi_max,
+                FLOAT chi_min, FLOAT chi_max,
+                FLOAT omega_min, FLOAT omega_max)
+        { 
+            FLOAT phi_mid = (phi_min + phi_max) / 2;
+            FLOAT chi_mid = (chi_min + chi_max) / 2;
+            FLOAT omega_mid = (omega_min + omega_max) / 2;
+            FLOAT fitness =  absolute(chi - chi_mid) + absolute(omega - omega_mid) + absolute(phi - phi_mid)/100.0;
+            // Big penalties for being out of the range
+            if (phi < phi_min || phi > phi_max) fitness += 10;
+            if (chi < chi_min || chi > chi_max) fitness += 10;
+            if (omega < omega_min || omega > omega_max) fitness += 10;
+            return;
         }
         """
 
@@ -787,7 +799,7 @@ class LimitedGoniometer(Goniometer):
             if (phi < -PI) phi += 2*PI;
             if (chi < -PI) chi += 2*PI;
             if (omega < -PI) omega += 2*PI;
-            fitness = fitness_function(phi, chi, omega);
+            fitness = fitness_function(phi, chi, omega, phi_min, phi_max, chi_min, chi_max, omega_min, omega_max);
             fitnesses.append(fitness);
             phi_list.append(phi);
             chi_list.append(chi);
@@ -803,9 +815,16 @@ class LimitedGoniometer(Goniometer):
         chi_list = []
         phi_list = []
         omega_list = []
+        phi_min = self.gonio_angles[0].random_range[0]
+        phi_max = self.gonio_angles[0].random_range[1]
+        chi_min = self.gonio_angles[1].random_range[0]
[…]

### 2010-12-16 gui/reflection_placer.py
Fixes #1988: Better reflection placer when limiting angles in a goniometer.

@@ -539,8 +539,8 @@ class FrameReflectionPlacer(wx.Frame):
             Group(Label("... or use the mouse to set the position by clicking below ...")),
             )
 
+            # Item("brute_search", label="Use brute-force search")
 
-            #Item("brute_search", label="Use brute-force search")
             
         viewBottom = View( \
             Group(Label("Measurement requires the following sample orientation:")),

### 2010-12-16 model/goniometer.py

@@ -447,7 +447,7 @@ class Goniometer(HasTraits):
             ub_matrix: for the current sample and mounting. B converts hkl to q-vector, and U orients.
             starting_angles: optional, list of angles from which to start the search.
                 e.g. we want the beam to move a few degrees away from this starting sample orientation.
-            search_method: option for the search method
+            search_method: option for the search method.
 
         Returns:
             angles: list of angles found
@@ -553,35 +553,92 @@ class Goniometer(HasTraits):
 class LimitedGoniometer(Goniometer):
     """Class for goniometers where there are limits to the range of phi, chi, and/or omega
     values. Holds general-purpose code for finding valid sample orientations."""
+    
+    name = String("Limited Goniometer")
+    description = String("Goniometer with 3 degrees of freedom but restrictions to allowed angles.")
 
     #-------------------------------------------------------------------------
     def __init__(self, wavelength_control=False, *args, **kwargs):
         Goniometer.__init__(self, wavelength_control, *args, **kwargs)
 
     #-------------------------------------------------------------------------
+    def are_angles_allowed(self, angles, return_reason=False):
+        """Calculate whether the given angles can be reached with the goniometer.
+        Will be overridden by subclasses as needed.
+
+        Parameters:
+            angles: list of angles of the goniometer (typically phi, chi, omega).
+            return_reason: do you want to return the reason (as string)?
+
+        Returns:
+            allowed: True if the angle provided can be reached.
+            reason: string describing the reason they cannot be reached. Empty if the angles are allowed.
+        """
+        ret = True
+        reason = ""
+
+        for i in xrange(len(angles)):
+            #@type AngleInfo ai
+            ai = self.gonio_angles[i]
+            angle = angles[i]
+            if angle < ai.random_range[0] or angle > ai.random_range[1]:
+                ret = False
+                reason += ai.name + " outside limits."
+
+        #The base goniometer can reach all angles.
+        if return_reason:
+            return (ret, reason)
+        else:
+            return ret
+
+
+    #-------------------------------------------------------------------------
     def get_fitness_function_c_code(self):
         #C code for the fitness of phi,chi, omega.
[…]

### 2010-12-16 docs/CrystalPlan_paper/png2pdf.sh
Refs #182

@@ -1,0 +1,3 @@
+# find . -name "*.png" -exec convert {} `basename {} .png`.pdf \;
+find . -name "*.png" -exec convert {} {}.pdf \;
+rename 's/\.png\.pdf/\.pdf/' *.pdf

### 2010-12-16 docs/CrystalPlan_paper/paper.tex
Fixes #186: Improved dialog for advanced goniometer angle editing.
Refs #182

@@
 
+\newif\ifdograph
+% \dographtrue % DO keep the graphs.
+\dographfalse % DONT put graphs - makes things faster
+
+\usepackage{ifpdf}
+
 \usepackage{graphicx}
+\DeclareGraphicsExtensions{.eps,.ps,.bmp,.tif,.tiff,.tga}  
+\graphicspath{{eps/}}
+
 \usepackage{gensymb} %General symbols
@@
 
-%\usepackage[pdftex]{graphicx}
-%\DeclareGraphicsExtensions{.pdf,.png,.jpg}
-
 %quick angstrom define
 \newcommand{\ang}{$\mathrm{\AA} $}
+%\newcommand{\th}[0]{\superscript{th}}
 
@@
 CrystalPlan will greatly increase the scientific productivity and throughput of
-TOPAZ and other beamlines. More users will be able to measure more samples in
-less time.
+TOPAZ and other beamlines. Users will be able to measure more
+samples or more fully explore parameters such as temperature and pressure.
 
@@
   in volume and for single peaks.
-  \item Take crystal symmetries into account.
+  \item Take crystal Bravais lattice symmetries into account.
   \item Optimize the experiment plan automatically.
+  \item Compare predicted and actual measurement of peaks.
 \end{itemize}
@@
 Each class inheriting from Goniometer can convert
-the desired phi, chi, omega angles to the internal motor positions it requires. The API allows
+the desired $\phi$, $\chi$, $\omega$ angles to the internal motor positions it requires. The API allows
 for limited goniometers: for example, TOPAZ is currently equipped with a
@@
 3D matrix is also specified (in \ang$^{-1}$), which determines the number of
-points to be modeled. Available machine memory limits the resolution that can be achieved in
-practice.
+points (voxels or volumetric pixels) to be modeled. Available machine memory
+limits the resolution that can be achieved in practice.
 
 Each point of the reciprocal space 3D matrix holds an integer representing the 
-number of times that voxel of q-space has been measured – this is the total
-coverage matrix. To speed up the calculation of the total, a coverage matrix for
-each sample orientation is calculated and saved in memory. This means that the
-total coverage matrix can be computed by simply adding multiple 3D matrices
-corresponding to each sample orientation.
+number of times that voxel of {\bf Q}-space (with {\bf Q} defined as $\frac{2\pi}{d}$) has
+been measured -- this is the total coverage matrix. To speed up the
+calculation of the total, a coverage matrix for each sample orientation is 
+calculated and saved in memory. This means that the total coverage matrix can
+be computed by simply adding multiple 3D matrices corresponding to each sample orientation.
 
@@
 For each sample orientation, a rotation matrix combining the sample's mounting
-orientation (the U matrix loaded from ISAW) and the goniometer rotation
+orientation (the U matrix loaded from ISAW) and the goniometer rotation matrices
 ($\Phi$, $\chi$, $\Omega$) is computed. The coverage in reciprocal space is
@@
 of the detectors and/or source. The scattered beam direction and the two possible incident beam vectors
-are used to to find two $q$ vectors, which point in the same direction. At the
-high wavelength we find $q_{min}$, and the low wavelength limit scattering off
-of $q_{max}$. Since the instrument has a wide bandwidth, we fill in the 3D q-space
-matrix by filling all points from $q_{min}$ to the $q_{max}$, for all pixels in
-the face of the detector. This quickly defines a volume of coverage, as shown
-schematically in two dimensions in Figure \ref{fig:ewald_fig}.
+are used to to find two $\overrightarrow q$ vectors; these point in the
+same direction. At the high wavelength we find $\overrightarrow q_{min}$, and
+the low wavelength limit scattering off of $\overrightarrow q_{max}$. Since the
+instrument has a wide bandwidth, we fill in the 3D {\bf Q}-space matrix by
+filling all points from $\overrightarrow q_{min}$ to the 
+$\overrightarrow q_{max}$, for all pixels in the face of the detector.
+This quickly defines a volume of coverage, as shown schematically in
+ two dimensions in Figure \ref{fig:ewald_fig}.
 
@@
 also be entered. This information is then used to calculate the increased
-measurement redundancy. A list of $n$ $3\times3$ multiplication matrices is
-generated based on the point group; any $hkl$ vector can be multiplied by
+measurement redundancy \cite{giacovazzo92}. A list of $n$ $3\times3$
+multiplication matrices is generated based on the point group; any $hkl$ vector can be multiplied by
 these matrices to find the $n$ equivalent $hkl$ reflections; for example, for
@@
 For the single reflections, each $hkl$ is tracked separately, but a primary
-$hkl$ is found that points to a list of equivalent $hkl$'s. The GUI can then display
-either all $hkl$, ignoring crystal symmetry; or combine them to display only part of
-the reflections.
+$hkl$ is found that points to a list of equivalent $hkl$'s. The GUI can
+then display either all $hkl$, ignoring crystal symmetry; or combine them to
+display only part of the reflections. Primary $hkl$'s
+are picked so that they make a contiguous block.
 
-In volume coverage mode, a 4D matrix is generated where each voxel has a list
-with $n$ entries, each pointing to a corresponding equivalent voxels. This
-allows CrystalPlan to quickly compute a volume coverage map that takes into
-account crystal symmetry: at each voxel, all $n$ voxels are checked to
-see if any of them were measured.
+In volume coverage mode, a 4D array is generated where the first 3 dimensions
+correspond to {\bf Q}-space, and the $4^{th}$ dimension is a list of $n$
+entries. Each entry in the $4^{th}$ dimension points to a voxel that is an
+equivalent due to crystal symmetry. This allows CrystalPlan to quickly compute a volume coverage
+map that takes into account crystal symmetry: at each voxel, all $n$ voxels are
+checked to see if any of them were measured.
 
@@
 user-specified coverage criterion (for example, the user may ask for $85\%$ of
-peaks to be measured, or for $90\%$ of q-volume).
+peaks to be measured, or for $90\%$ of {\bf Q}-volume).
 
-Genetic algorithms have been covered in the literature before, see e.g ref
-[INSERT REFERENCE HERE]. In this application, the genes consist of goniometer
-angles. For instance, if a goniometer has freedom in 3 angles and the user
+Genetic algorithms have been covered in the literature before, see e.g
+\cite{Goldberg89}. In this application, the genes
+consist of goniometer angles. For instance, if a goniometer has freedom in 3 angles and the user
 requests a 10 different sample orientations, then each individual has 30 genes.
@@
 the difficulty of the problem and the computing hardware. Given that at a
-neutron facility, a measurement at a single sample orientation may take a few
+neutron facility, a measurement of a single orientation may take a few
 hours, this feature can result in significant beam-time savings.
@@
 
-To this end, CrystalPlan was written in Python, a powerful, open, cross-platform
-scripting language. Python was chosen for its ease of cross-platform deployment, and for
+To this end, CrystalPlan was written in Python \cite{python}, a powerful, open,
+cross-platform scripting language. Python was chosen for its ease of cross-platform deployment, and for
 its attractive language features. The numpy and scipy libraries (two
 open-source, commonly installed Python libraries) were used for many of the
-calculations. Some of the critical calculations were also written with inline C
-code, which speed up some calculations by up to $400\times$ as compared with
+calculations \cite{numpy}. Some of the critical calculations were also written
+with inline C code, which speed up some calculations by up to $400\times$ as compared with
 the pure Python version (which can still be used, in case of incompatibilities).
@@
 CrystalPlan, since the program is meant to be used interactively as a guide to
-the experimenter. The user interface was written using wxPython, a mature
-cross-platform GUI toolkit, as well as the Enthought Traits GUI. Some of the 2D
-plots use matplotlib, and the 3D visualization elements use the Mayavi2 toolkit
-for Python, itself based on VTK. All of these libraries are open-source.
+the experimenter. The user interface was written using wxPython
+\cite{wxPython}, a mature cross-platform GUI toolkit, as well as the Enthought
+Traits GUI. Some of the 2D plots use matplotlib \cite{matplotlib},
+and the 3D visualization elements use the Mayavi2 toolkit for Python
+\cite{mayavi}, itself based on VTK \cite{VTK}. All of these libraries are
+open-source.
 
@@
 \begin{itemize}
-  	\item Q-Space tab: define the volume of interest (minimum d-spacing) and
-resolution of reciprocal space. 
+  	\item {\bf Q}-Space tab: define the volume of interest (minimum d-spacing)
+  	and resolution of reciprocal space. 
  	\item Detectors tab: load instrument detector geometry
@@
 complete, the plan can be sent to the SNS data acquisition computers in order to
-begin a run.
+begin a run \cite{SNSDAS}.
 
@@
 Figure \ref{fig:volume_view}a) shows a screenshot of the reciprocal space
-coverage 3D interface. The 3D view can represent q-space in two ways: volume coverage, or single-peak
+coverage 3D interface. The 3D view can represent {\bf Q}-space in two ways: volume coverage, or single-peak
 coverage. The view shown can be rotated, translated or zoomed using the mouse or
@@
 
-%\includegraphics[bb=0in 0in 8in 6in, scale=0.20]{ewald_fig.pdf}
+\ifdograph
   \includegraphics[width=3.5in]{ewald_fig.ps}
+\fi
 
@@
 top of the main window.}
-\includegraphics{main_gui.eps}
+\ifdograph
+  \includegraphics{main_gui}
+\fi
 \label{fig:screenshots}
@@
 symmetry), coverage increases to 65\%. }
-\includegraphics{volume_nosym.eps}
+\ifdograph
+  \includegraphics{volume_nosym}
+\fi
 a)
-\includegraphics{volume_sym.eps}
+\ifdograph
+  \includegraphics{volume_sym}
+\fi
 b)
@@
 89\%, and redundancy is also increased. }
-\includegraphics{refl_nosym.eps}
+\ifdograph
+  \includegraphics{refl_nosym} 
+\fi
 a)
-\includegraphics{refl_sym.eps}
+\ifdograph
+  \includegraphics{refl_sym}
+\fi
 b)
@@
 single $hkl$.}
-\includegraphics[width=2.5in]{single_reflection_2.eps}
+\ifdograph
+  \includegraphics[width=2.5in]{single_reflection_3}
+\fi
+
 \label{fig:single_reflection} 
@@
 statistics.}
-\includegraphics[width=3.5in]{Plan8List.eps}
+\ifdograph
+  \includegraphics[width=3.5in]{Plan8List}
+\fi
 a)
-\includegraphics[width=3.5in]{NatCoverage.eps}
+\ifdograph
+  \includegraphics[width=3.5in]{NatCoverage}
+\fi
 b)

### 2010-12-16 docs/CrystalPlan_paper/paper.tex
Refs #182. All Pete's comments are in.

@@
 open-source, commonly installed Python libraries) were used for many of the
-calculations \cite{numpy}. Some of the critical calculations were also written
-with inline C code, which speed up some calculations by up to $400\times$ as compared with
+calculations \cite{numpy,scipy}. Some of the critical calculations were
+also written with inline C code, which speed up some calculations by up to $400\times$ as compared with
 the pure Python version (which can still be used, in case of incompatibilities).
@@
 The calculation code of CrystalPlan is placed in a model layer and kept
-completely separate from eventual GUI elements. This makes it possible to run
-calculations using Python scripts, if desired. Unit tests for each part of the
-calculations are used to ensure consistent results on different platforms and
-systems.
+completely separate from eventual GUI elements, using a Model-View-Controller
+architecture \cite{Gamma95}. This makes it possible to run calculations using
+Python scripts, if desired. Unit tests for each part of the calculations are used to ensure
+consistent results on different platforms and  systems.
 
@@
 RMS difference between the predicted and actual measurement wavelength was only
-0.016 \ang. Inspecting the detector position offsets showed that the main
-contributor to this error was a slight misalignment of the goniometer, which
-CrystalPlan can help diagnose.
+0.016 \ang. The offsets in the peak position were plotted, which
+helped show that the main contributor to this error was a slight misalignment of
+the goniometer.
 
@@
 efficient experiment plans for wide bandwidth time-of-flight Laue
-diffractometers. This program will help maximize scientific productivity at SNS
+diffractometers. This program will help increase scientific productivity at SNS
 beamlines such as TOPAZ.

### 2010-12-16 docs/CrystalPlan_paper/paper.tex
Refs #182. All Pete's comments are in.

@@
 \newif\ifdograph
-% \dographtrue % DO keep the graphs.
-\dographfalse % DONT put graphs - makes things faster
+\dographtrue % DO keep the graphs.
+%\dographfalse % DONT put graphs - makes things faster
 
@@
 
-\cauthor[a]{Janik}{Zikovsky}{zikovskyjl@ornl.gov}{address if different from
+\cauthor[]{Janik}{Zikovsky}{zikovskyjl@ornl.gov}{address if different from
 \aff}
-\author[a]{Peter}{Peterson}
-\author[a]{Xiaoping}{Wang}
-\author[a]{Matthew}{Frost}
-\author[a]{Christina}{Hoffmann}
+\author[]{Peter}{Peterson}
+\author[]{Xiaoping}{Wang}
+\author[]{Matthew}{Frost}
+\author[]{Christina}{Hoffmann}
 
-\aff[a]{Spallation Neutron Source, Oak Ridge National Laboratory, P.O. Box 2008
+\aff[]{Spallation Neutron Source, Oak Ridge National Laboratory, P.O. Box 2008
 MS-6477, Oak Ridge, TN 37831-6477
@@
 \ifdograph
-  \includegraphics[width=2.5in]{single_reflection_3}
+  \includegraphics[width=3in]{single_reflection_3}
 \fi

### 2010-12-17 docs/CrystalPlan_paper/paper.tex
Refs #182: Paper

@@
 
-From an initial data set, the data analysis program ISAW \cite{Tao2006} is used
+From an initial data set, the data analysis program ISAW \cite{Mikkelson05} is used
 to calculate the the sample mounting orientation UB matrix, which is saved to a text file along with the lattice

### 2010-12-30 docs/CrystalPlan_paper/paper.tex
Refs #182: Christina's changes. Should be the last draft before submitting.

--- 2010-12-16 docs/CrystalPlan_paper/paper.tex
@@
 \begin{abstract}
-Beam time at large x-ray and neutron scattering facilities is always at a premium.
-The CrystalPlan program can calculate the data coverage of a crystal in reciprocal
-space in a single-crystal diffraction time-of-flight experiment. CrystalPlan can 
-help a user build an experiment plan that will acquire the most data possible, 
-with sufficient coverage but limited redundancy, therefore increasing scientific 
-productivity. 
+Beam time at large user program based x-ray and neutron scattering facilities is
+in high demand and always at a premium. CrystalPlan, a highly efficient
+experiment planning software has been developed to maximize the use of available
+beamtime per sample per experiment. 
+This program can calculate and optimize the data coverage of a crystal in
+reciprocal space in a single-crystal diffraction time-of-flight experiment.
+CrystalPlan can help a user build an experiment plan that will acquire the most
+data possible, with sufficient coverage but limited redundancy, therefore
+increasing scientific productivity. 
 A user friendly GUI including a 3D viewer, an automated coverage optimizer,
@@
 The TOPAZ beamline at the Spallation Neutron Source (SNS) at Oak Ridge National
-Lab (ORNL), currently finishing commissioning, is a neutron time-of-flight
-Laue instrument with high wavelength resolution. It is designed to acquire
-single-crystal diffraction data of small-to-moderate sized unit cells with  high
-throughput. The final design calls for an array of 48 detectors covering a large
-fraction of $2\pi$ steradians. However as of this writing, only 14 of the 48
-detectors are installed, giving only a fraction of $2\pi$ steradian coverage. 
-In order to have sufficient data to perform a fit, an experimenter will typically want
-to measure $ > 85\%$ of the peaks within a certain range of d-spacings. 
-However, wide wavelength band combined with TOPAZ's complex
-geometry makes it very difficult to come up with an effective set of sample
-orientations that will cover all the required peaks without unnecessary redundancy.
+Laboratory (ORNL), currently finishing commissioning, is a neutron
+time-of-flight Laue diffractometer with high wavelength (time) resolution. It
+is designed to acquire single-crystal diffraction data of small-to-moderate sized unit cells with  high
+throughput. The final design calls for an array of 48 detectors covering three
+quarters of $2\pi$ steradians. However as of this writing, only 14 of the 48
+detectors are installed, giving only a quarter of $2\pi$ steradian coverage. In
+order to have sufficient data for a reliable structure solution and
+refinement, an experimenter will typically want to measure $ > 85\%$ of the
+peaks within a certain range of d-spacings. However, wide wavelength band
+combined with TOPAZ's complex geometry makes it very difficult to come up with
+an effective set of sample orientations that will cover all the required peaks
+without unnecessary redundancy.
 
@@
 CrysAlis Pro from Oxford Diffraction; COSMO from Bruker AXS) were designed for
-X-ray experiments where a single  incident wavelength and a large number of
-sample orientations or sweeps are used to collect data. 
-These programs are not suitable for TOPAZ, since as a time-of-flight instrument, 
-it is able to collect data from a wide bandwidth of
-incident neutrons, using a stationary crystal. 
-Additionally, this means that relatively few sample orientations are needed in
-order to complete a measurement: whereas an X-ray measurement might have
-hundreds or thousands of individual frames, a TOPAZ run may have as few as 10.
+X-ray experiments where the crystal is rotated stepwise through the Bragg
+diffraction condition using a single incident wavelength, 
+resulting in a large number of sample orientations.
+These programs are not suitable for TOPAZ, since as a time-of-flight instrument
+(time-resolved Laue), it is collecting data from a wide
+bandwidth of incident neutrons, contracting the Ewald sphere through the
+reflection condition thus while keeping a crystal stationary, as described in
+[!!!! CC WILSON REFERENCE].
+ Additionally, this means that relatively few sample
+orientations are needed in order to complete a measurement: whereas an X-ray measurement might have
+hundreds or thousands of individual frames, a TOPAZ run may have as few as 10
+to cover the same volume of reciprocal space.
 
@@
 planning their experiments. By quickly making an experimental plan that will be
-sufficient for the needs of the scientist (enough peaks measured), but without
-spending time needlessly measuring the same region in reciprocal space,
-CrystalPlan will greatly increase the scientific productivity and throughput of
-TOPAZ and other beamlines. Users will be able to measure more
-samples or more fully explore parameters such as temperature and pressure.
+tailored to the needs of the scientist (enough peaks measured for a structure
+analysis), but without spending time collecting redundant data, CrystalPlan will
+greatly increase the scientific productivity and throughput of TOPAZ and other
+beamlines. Users will be able to measure samples highly efficiently, allowing
+them to explore parameters such as temperature and pressure more fully.
 
@@
 \begin{itemize}
-  \item Load an instrument geometry.
+  \item Load an instrument geometry file.
   \item Calculate the orientation of a crystal at different goniometer settings.
@@
   in volume and for single peaks.
-  \item Take crystal Bravais lattice symmetries into account.
+  \item Take crystal Bravais lattice symmetries and the higher redundancy of
+  higher space groups into account.
   \item Optimize the experiment plan automatically.
@@
 
-CrystalPlan supports area detectors with flat faces and arbitrary rectangular
-dimensions; this handles the detector types in use at SNS (Anger cameras and
-8-pack tube detectors), but the software could easily be extended for more
-complicated detector shapes as needed. The detector configuration is loaded from
+CrystalPlan currently supports area detectors with flat faces and arbitrary
+rectangular dimensions; this reflects the detector types in use at SNS (Anger
+cameras and 8-pack $^3$He tube detectors), but the software could easily be
+extended for more complicated detector shapes as needed. The detector configuration is loaded from
 a text file containing the location, size, orientation and number of pixels of
-each detector. Other settings to be set for an instrument include the
+each detector. Other settings to be adjusted for an instrument include the
 detectable range of neutron/x-ray wavelength.
@@
 
-CrystalPlan is designed to be able to use a variety of sample orientation
-goniometers. A base Goniometer class is subclassed for specific types of
+CrystalPlan can accomodate a variety of sample
+orientation goniometers. A base Goniometer class is subclassed for specific types of
 goniometers, allowing a unified interface. This base class acts as a
 \emph{universal goniometer},
-which uses the conventional goniometer angles of
+which uses the conventional Eulerian goniometer angles of
 $\phi$, $\chi$, $\omega$ as defined in \cite{busing67}.
-Each class inheriting from Goniometer can convert
-the desired $\phi$, $\chi$, $\omega$ angles to the internal motor positions it requires. The API allows
-for limited goniometers: for example, TOPAZ is currently equipped with a
-goniometer where chi is fixed at 45 degrees. CrystalPlan handles this by
-limiting the available sample orientation angles accessible to the user.
+Each class inheriting from the base Goniometer can convert
+the desired $\phi$, $\chi$, $\omega$ angles to the internal motor positions it requires. 
+The class hierarchy allows for goniometer limitations: for example, TOPAZ is
+currently equipped with a goniometer where $\chi$ is fixed at 45 degrees.
+CrystalPlan handles this by limiting the sample orientation angles accessible to the user to $\phi$ and
+$\omega$.
 
-From an initial data set, the data analysis program ISAW \cite{Tao2006} is used
-to calculate the the sample mounting orientation UB matrix, which is saved to a text file along with the lattice
-parameters. CrystalPlan can load this UB matrix file and will use that UB matrix
-in coverage calculations.
+The sample unit cell and orientation is typically determined by the data
+collection/reduction software. In our case, from an initial data set, the data
+analysis program ISAW (Integrated Scattering Analysis
+Workbench \cite{Mikkelson05}) provides us with the unit cell of the sample (B
+matrix) and the sample mounting orientation (U matrix).
+The combined UB matrix is then saved to a
+text file along with the lattice parameters. 
+CrystalPlan can load this UB matrix file and uses it for reciprocal coverage
+calculations and optimization.
 
@@
 
-The reciprocal space to be measured is divided into a cubic 3D matrix, the size
-of which is determined by the user, who specifies the minimum d-spacing
-($d_{min}$, in \ang) that he or she is interested in. The resolution of the
-3D matrix is also specified (in \ang$^{-1}$), which determines the number of
-points (voxels or volumetric pixels) to be modeled. Available machine memory
+The reciprocal space to be measured is divided into a cubic 3D matrix 
+of {\bf Q}-space (with {\bf Q} defined as $\frac{2\pi}{d}$). 
+The user specifies the minimum d-spacing ($d_{min}$, in \ang) 
+that he or she is interested in, defining the limit
+$Q_{max} = \frac{2\pi}{d_{min}}$ of the modeled volume. There is no lower limit
+on {\bf Q}.
+
+The resolution of the 3D matrix is also specified
+(in \ang$^{-1}$), which determines the number of points (voxels or volumetric pixels) to be modeled. Available machine memory
 limits the resolution that can be achieved in practice.
@@
 Each point of the reciprocal space 3D matrix holds an integer representing the 
-number of times that voxel of {\bf Q}-space (with {\bf Q} defined as $\frac{2\pi}{d}$) has
+number of times that voxel of {\bf Q}-space has
 been measured -- this is the total coverage matrix. To speed up the
 calculation of the total, a coverage matrix for each sample orientation is 
-calculated and saved in memory. This means that the total coverage matrix can
-be computed by simply adding multiple 3D matrices corresponding to each sample orientation.
+calculated and saved in memory and the total coverage
+matrix can be computed by simply adding multiple 3D matrices corresponding to each sample orientation.
 
@@
 of the detectors and/or source. The scattered beam direction and the two possible incident beam vectors
-are used to to find two $\overrightarrow q$ vectors; these point in the
-same direction. At the high wavelength we find $\overrightarrow q_{min}$, and
-the low wavelength limit scattering off of $\overrightarrow q_{max}$. Since the
-instrument has a wide bandwidth, we fill in the 3D {\bf Q}-space matrix by
-filling all points from $\overrightarrow q_{min}$ to the 
+are used to find two $\overrightarrow q$ vectors; these point in the
+same direction. At the high wavelength limit we find $\overrightarrow q_{min}$,
+and the low wavelength limit is scattering off of $\overrightarrow q_{max}$.
+Since the instrument has a wavelength bandwidth, we fill in the 3D {\bf Q}-space
+matrix by filling all points from $\overrightarrow q_{min}$ to the 
 $\overrightarrow q_{max}$, for all pixels in the face of the detector.
-This quickly defines a volume of coverage, as shown schematically in
+A typical range for TOPAZ is 0.5 to 3.6 \ang, a 3 \ang\ wavelength band.
+This readily defines a volume of coverage, as shown schematically in
 two dimensions in Figure \ref{fig:ewald_fig}.
@@
 To reduce memory usage yet allow for quickly trying different detector
-configurations, the 3D matrix of each sample orientation is saved a 32- or
+configurations, the 3D matrix of each sample orientation is saved as a 32- or
 64-bit integer, with each bit representing a particular detector. A simple
@@
 
-Each $hkl$ peak is a separate object that holds a list of predicted
-measurements, including which detector measured it, at what position on the
+Each $hkl$ peak is a separate object that holds all predicted
+recordings, including which detector measured it, at what (x,y) position on the
 detector face, at what wavelength, and for which sample orientation. These data
-will be used to  graphically display coverage and calculate statistics such as
+is used to  graphically display coverage and calculate statistics such as
 redundancy.
@@
 \subsection{Crystal Symmetry}
-
-When entering the lattice parameters, the crystal's point group symmetry can
-also be entered. This information is then used to calculate the increased
-measurement redundancy \cite{giacovazzo92}. A list of $n$ $3\times3$
-multiplication matrices is generated based on the point group; any $hkl$ vector can be multiplied by
+\label{sec:symm}
+The crystal's point group symmetry can
+be entered along with its lattice parameters.
+This information is then used to
+calculate the increased measurement redundancy for higher symmetries
+\cite{giacovazzo92}. 
+A list of $n$ $3\times3$ multiplication matrices is generated based on the point group; 
+all $hkl$ vectors are multiplied by
 these matrices to find the $n$ equivalent $hkl$ reflections; for example, for
@@
 
-For the single reflections, each $hkl$ is tracked separately, but a primary
-$hkl$ is found that points to a list of equivalent $hkl$'s. The GUI can
-then display either all $hkl$, ignoring crystal symmetry; or combine them to
-display only part of the reflections. Primary $hkl$'s
-are picked so that they make a contiguous block.
+Each reflection is tracked separately, but a primary
+$hkl$ is identified that points to a list of equivalent $hkl$'s when the
+crystal symmetry is known.
+The interactive GUI displays either all $hkl$, ignoring crystal symmetry; or
+combines them to display only the primary $hkl$'s, and calculates the
+redundancy by summing all measurements on all equivalent $hkl$'s.
 
@@
 equivalent due to crystal symmetry. This allows CrystalPlan to quickly compute a volume coverage
-map that takes into account crystal symmetry: at each voxel, all $n$ voxels are
-checked to see if any of them were measured.
+map that takes into account crystal symmetry and the resulting redundancy: at
+each voxel, all $n$ equivalent voxels are checked to see if any of them were
+measured.
 
@@
 CrystalPlan allows the user to manually enter lists of sample orientations to
-simulate; however, finding an optimal plan by trial and error would be very time
-consuming. To make the process easier, CrystalPlan includes an Automatic
+simulate and graphically display the coverage in reciprocal space;
+however, finding an optimal plan by trial and error is in general very
+time consuming and only feasible for an experienced crytallographer.
+To make the process more efficient, CrystalPlan includes an Automatic
 Coverage Optimizer, which takes as its inputs the desired number of sample
@@
 user-specified coverage criterion (for example, the user may ask for $85\%$ of 
-peaks to be measured, or for $90\%$ of {\bf Q}-volume).
+Bragg reflections to be measured, or for $90\%$ of {\bf Q}-volume).
 
-Genetic algorithms have been covered in the literature before, see e.g
+Genetic algorithms have been covered in the literature before, see e.g.
 \cite{Goldberg89}. In this application, the genes
 consist of goniometer angles. For instance, if a goniometer has freedom in 3 angles and the user
-requests a 10 different sample orientations, then each individual has 30 genes.
+requests 10 different sample orientations, then each individual has 30 genes.
 The initial population is created randomly and is made to evolve using mutations
 and crossover. Individuals whose goniometer angles turn out to be impossible to
-reach are eliminated. For each individual set of sample orientations, the
-coverage statistics are computed (taking crystal symmetry into account if desired), and this is used as the fitness of the given individual in the normal genetic
-algorithm process.
+reach or beyond set limits are eliminated. For each individual set of sample
+orientations, the coverage statistics are computed (taking crystal symmetry into
+account if desired), and this is used as the fitness of the given individual in
+the normal genetic algorithm process. The best individuals are saved after each
+cycle. The simulation has succeeded once the desired value is reached or
+exceeded or the maximum number of cycles was calculated.
 
@@
 
-The program needs to be cross-platform, and run equally well on Linux, Mac OS
-and Windows, although Linux is our primary target operating system. Another
+The program was designed to be cross-platform, and run equally well on Linux,
+Mac OS and Windows, although Linux is our primary target operating system. Another
 design goal was to use of open-source software and toolkits as much as possible.
@@
 open-source, commonly installed Python libraries) were used for many of the
-calculations \cite{numpy,scipy}. Some of the critical calculations were
-also written with inline C code, which speed up some calculations by up to $400\times$ as compared with
-the pure Python version (which can still be used, in case of incompatibilities).
+calculations \cite{numpy,scipy}. The most time-critical calculations
+were also written with inline C code, which increases the computation speed by
+up to $400\times$ as compared with the pure Python version (which can still be
+used, in case of incompatibilities).
 
@@
 The calculation code of CrystalPlan is placed in a model layer and kept
-completely separate from eventual GUI elements, using a Model-View-Controller
+completely separate from eventual GUI elements, using a Model/View-Controller
 architecture \cite{Gamma95}. This makes it possible to run calculations using
@@
 CrystalPlan, since the program is meant to be used interactively as a guide to
-the experimenter. The user interface was written using wxPython
+the experimenter whether new or experienced. This requires intuitive GUI
+interfaces and workflows.
+The user interface was written using wxPython
 \cite{wxPython}, a mature cross-platform GUI toolkit, as well as the Enthought
-Traits GUI. Some of the 2D plots use matplotlib \cite{matplotlib},
-and the 3D visualization elements use the Mayavi2 toolkit for Python
-\cite{mayavi}, itself based on VTK \cite{VTK}. All of these libraries are
-open-source.
+Traits GUI. Some of the 2D plots use matplotlib \cite{matplotlib},  and the 3D
+visualization elements use the Mayavi2 toolkit for Python \cite{mayavi}, itself
+based on VTK \cite{VTK}. All of these libraries are  open-source.
 
@@
 	\item Goniometer tab: to choose the
-goniometer and degrees of orientation freedom of the instrument.
+goniometer and degrees of orientation freedom of the sample.
   	\item Sample tab: enter or load the crystal's lattice parameters and UB
@@
 delete, or temporarily disable them to see the effect on coverage. Once
-complete, the plan can be sent to the SNS data acquisition computers in order to
-begin a run \cite{SNSDAS}.
+complete, the plan can be sent to the SNS/TOPAZ data acquisition computers in
+order to begin a run \cite{SNSDAS}.
 
@@
 isosurfaces; or the coverage can be inverted, showing non-measured volumes only;
-or sliced between user-selectable $q_{min}$ and $q_{max}$.
+or slices between user-selectable $q_{min}$ and $q_{max}$.
 
@@
 
-Both views can take into crystal symmetry in their display using the techniques
-described previously. 
+Both views can take into account crystal symmetry in their display using the
+techniques described previously in Section \ref{sec:symm}.
 The volume coverage (Figures \ref{fig:volume_view}b) fills in the sphere of
@@
 
-Finally, both views display quick coverage statistics in the lower right
+Finally, both views display coverage statistics in the lower right
 corner, showing the percentage of volume (or of the number of peaks) measured,
@@
 position on a detector by automatically finding the necessary sample orientation
-angles. This is used, for example, to center important peaks in the most
-sensitive part of a detector.
+angles. This is used, for example, to center important peaks or diffraction
+directions in the most sensitive part of a detector.
 
@@
 orientations (Figure \ref{fig:natrolite_gui}a) that covered 97\% of the peaks,
-taking into account the crystal symmetry (orthorombic or mmm) (Figure
-\ref{fig:natrolite_gui}b). The plan was left to run unattended overnight.
+taking into account the crystal Laue symmetry (orthorombic or mmm) (Figure
+\ref{fig:natrolite_gui}b). The eight orientations were run
+unattended overnight.
 
@@
 recently developped capabilities for real-time data analysis to monitor
-the  signal to noise ratio on particular peaks. When the signal becomes
-sufficient, the sample will automatically be moved to the next sample
-orientation, further minimizing data acquisition time. Combined with advanced
-sample changers coming on-line at TOPAZ, this will allow for extremely
-efficient, unattended data acquisition.
+the  signal to noise ratio on particular peaks. When the signal intensity
+becomes sufficient, the sample will automatically be moved to the next sample
+orientation, further adjusting data acquisition time for each specific sample.
+Combined with advanced sample changers coming on-line at TOPAZ, this will allow
+for extremely efficient, unattended and automated data acquisition.
 

### 2010-12-30 docs/CrystalPlan_paper/paper.tex
Refs #182: Last couple of references added.

@@
 reflection condition thus while keeping a crystal stationary, as described in
-[!!!! CC WILSON REFERENCE].
+\cite{Schultz94,Wilson00}.
  Additionally, this means that relatively few sample
@@
 \begin{itemize}
-  \item Load an instrument geometry file.
+  \item Load a geometry file for a specific instrument (using a generic,
+  flexible file format).
   \item Calculate the orientation of a crystal at different goniometer settings.

### 2011-01-28 docs/CrystalPlan_paper/paper.tex
Refs #182: Implemented Bryan's corrections.

@@
 and an option to reorient the crystal for the measurement 
-of selected hkls on specific detector positions 
+of selected $hkl$s on specific detector positions 
 are among its useful features. 
@@
 quarters of $2\pi$ steradians. However as of this writing, only 14 of the 48
-detectors are installed, giving only a quarter of $2\pi$ steradian coverage. In
-order to have sufficient data for a reliable structure solution and
+detectors are installed, giving only a quarter of $2\pi$ steradian coverage. 
+To have sufficient data for a reliable structure solution and
 refinement, an experimenter will typically want to measure $ > 85\%$ of the
-peaks within a certain range of d-spacings. However, wide wavelength band
-combined with TOPAZ's complex geometry makes it very difficult to come up with
+peaks within a certain range of d-spacings. However, TOPAZ's wide wavelength
+band combined with its complex geometry makes it difficult to come up with
 an effective set of sample orientations that will cover all the required peaks
@@
 bandwidth of incident neutrons, contracting the Ewald sphere through the
-reflection condition thus while keeping a crystal stationary, as described in
+reflection condition thus while keeping the crystal stationary, as described in
 \cite{Schultz94,Wilson00}.
  Additionally, this means that relatively few sample
-orientations are needed in order to complete a measurement: whereas an X-ray measurement might have
+orientations are needed to complete a measurement: whereas an X-ray measurement might have
 hundreds or thousands of individual frames, a TOPAZ run may have as few as 10
@@
 greatly increase the scientific productivity and throughput of TOPAZ and other
-beamlines. Users will be able to measure samples highly efficiently, allowing
-them to explore parameters such as temperature and pressure more fully.
+similar beamlines. Users will be able to measure samples highly efficiently,
+allowing them to explore parameters such as temperature and pressure more fully.
 
@@
   in volume and for single peaks.
-  \item Take crystal Bravais lattice symmetries and the higher redundancy of
-  higher space groups into account.
+  \item Take into account crystal Bravais lattice symmetries and the higher
+  redundancy of higher space groups.
   \item Optimize the experiment plan automatically.
@@
 
-\subsection{Reciprocal Space Representation.}
+\subsection{Reciprocal Space Representation}
 
@@
 detector face, at what wavelength, and for which sample orientation. These data
-is used to  graphically display coverage and calculate statistics such as
+are used to graphically display coverage and calculate statistics such as
 redundancy.
@@
 simulate and graphically display the coverage in reciprocal space;
-however, finding an optimal plan by trial and error is in general very
+however, finding an optimal plan by trial and error is in general 
 time consuming and only feasible for an experienced crytallographer.
@@
 
-As part of the TOPAZ beamline commissioning process, a Natrolite
+As part of the TOPAZ beamline commissioning process, a natrolite
 (\ce{Na2Al2Si3O10} $\cdot$ \ce{2H2O}) single crystal was measured on a fixed
@@
 positions on each detector face were compared for the 554 integrated peaks
-found. Very good agreement was found; the
-RMS difference between the predicted and actual measurement wavelength was only
-0.016 \ang. The offsets in the peak position were plotted, which
-helped show that the main contributor to this error was a slight misalignment of
-the goniometer.
+found. Good agreement was found; the RMS difference between the predicted and
+actual measurement wavelength was only 0.016 \ang. The offsets in the peak
+position were plotted, which helped show that the main contributor to this error
+was a slight misalignment of the goniometer.
 
@@
 CrystalPlan would interface with data analysis software such as ISAW and use
-recently developped capabilities for real-time data analysis to monitor
+recently developed capabilities for real-time data analysis to monitor
 the  signal to noise ratio on particular peaks. When the signal intensity

### 2011-01-28 docs/CrystalPlan_paper/paper.tex
Refs #182: Some of the reviewer's corrections.

@@
 \dographtrue % DO keep the graphs.
-%\dographfalse % DONT put graphs - makes things faster
+\dographfalse % DONT put graphs - makes things faster
 
@@
 increasing scientific productivity. 
-A user friendly GUI including a 3D viewer, an automated coverage optimizer,
-and an option to reorient the crystal for the measurement 
+A user friendly GUI (Graphical User Interface) including a 3D viewer, an
+automated coverage optimizer, and an option to reorient the crystal for the measurement 
 of selected $hkl$s on specific detector positions 
@@
 (time-resolved Laue), it is collecting data from a wide
-bandwidth of incident neutrons, contracting the Ewald sphere through the
-reflection condition thus while keeping the crystal stationary, as described in
+bandwidth of incident neutrons, thus contracting the Ewald sphere through the
+reflection condition while keeping the crystal stationary, as described in
 \cite{Schultz94,Wilson00}.
@@
 CrystalPlan can accomodate a variety of sample
-orientation goniometers. A base Goniometer class is subclassed for specific types of
-goniometers, allowing a unified interface. This base class acts as a
-\emph{universal goniometer},
+orientation goniometers. A base Goniometer class (in the object-oriented
+programming sense) is subclassed for specific types of goniometers, allowing a
+unified interface. This base class acts as a \emph{universal goniometer},
 which uses the conventional Eulerian goniometer angles of
@@
 analysis program ISAW (Integrated Scattering Analysis
-Workbench \cite{Mikkelson05}) provides us with the unit cell of the sample (B
-matrix) and the sample mounting orientation (U matrix).
+Workbench \cite{Mikkelson05}) provides us with the unit cell of the sample (via
+the B matrix) and the sample mounting orientation (U matrix).
 The combined UB matrix is then saved to a
@@
 same direction. At the high wavelength limit we find $\overrightarrow q_{min}$,
-and the low wavelength limit is scattering off of $\overrightarrow q_{max}$.
-Since the instrument has a wavelength bandwidth, we fill in the 3D {\bf Q}-space
+and the low wavelength limit corresponds to $\overrightarrow
+q_{max}$. Since the instrument has a wavelength bandwidth, we fill in the 3D {\bf Q}-space
 matrix by filling all points from $\overrightarrow q_{min}$ to the 
 $\overrightarrow q_{max}$, for all pixels in the face of the detector.
-A typical range for TOPAZ is 0.5 to 3.6 \ang, a 3 \ang\ wavelength band.
+A typical wavelength band for TOPAZ is from 0.5 to 3.6 \ang.
 This readily defines a volume of coverage, as shown schematically in
@@
 these matrices to find the $n$ equivalent $hkl$ reflections; for example, for
-orthorombic (mmm) symmetry, $n=8$; therefore 8 multiplication matrices are
+orthorhombic (mmm) symmetry, $n=8$; therefore 8 multiplication matrices are
 generated that multiply $hkl$ to give $\pm h$, $\pm k$, $\pm l$.
@@
 
-The program was designed to be cross-platform, and run equally well on Linux,
+The program was designed to be cross-platform, and to run equally well on Linux,
 Mac OS and Windows, although Linux is our primary target operating system. Another
-design goal was to use of open-source software and toolkits as much as possible.
+design goal was to use open-source software and toolkits as much as possible.
 
@@
 orientations (Figure \ref{fig:natrolite_gui}a) that covered 97\% of the peaks,
-taking into account the crystal Laue symmetry (orthorombic or mmm) (Figure
+taking into account the crystal Laue symmetry (orthorhombic or mmm) (Figure
 \ref{fig:natrolite_gui}b). The eight orientations were run
@@
 positions on each detector face were compared for the 554 integrated peaks
-found. Good agreement was found; the RMS difference between the predicted and
-actual measurement wavelength was only 0.016 \ang. The offsets in the peak
+found. Good agreement was found; the root-mean-square difference between the
+predicted and actual measurement wavelength was only 0.016 \ang. The offsets in the peak
 position were plotted, which helped show that the main contributor to this error
 was a slight misalignment of the goniometer.
-
-%, whereas the RMS of the detector position error was 7.6 mm.
 
@@
 \begin{figure}
-\caption{2D Ewald sphere construction showing reciprocal-space detector coverage
+\caption{2D Ewald construction showing reciprocal-space detector coverage
 for a wide-bandwidth Laue diffractometer. The smaller blue circle represents the
@@
 wedge, and projects an arc onto each sphere of radius $1/\lambda$. The
-red wedge between the two $\lambda$ is built by projecting straight lines between the
-$q_{min}$ and $q_{max}$ limits for each pixel on the detector face, giving the
+red wedge between the two $\lambda$ is built by drawing straight lines between
+the $q_{min}$ and $q_{max}$ limits for each pixel on the detector face, giving the
 volume in reciprocal space that is measured at a single sample oriention. }
@@
 \caption{Sample screenshot showing the 3D view of the predicted coverage of
-reciprocal space volume for 3 sample orientations and 14 detectors in TOPAZ, from 0.9 \ang\ to 4.0\ang:
-a) without crystal symmetry, 17\% coverage; b) with crystal symmetry (mmm 8-fold
+reciprocal space volume for 3 sample orientations and 14 detectors on TOPAZ,
+from 0.9 \ang\ to 4.0\ang: a) without crystal symmetry, 17\% coverage; b) with crystal symmetry (mmm 8-fold
 symmetry), coverage increases to 65\%. }
@@
 measurements. Each sphere represents an $hkl$ peak, color-coded to indicate the
-number of measurements (dark blue: 0, cyan: 1, green: 2, etc.). 
-a) without crystal symmetry, 2375 of 9481 reflections are measured, or 25\%.
-b) with crystal symmetry, 1276 of 1427 primary reflections are measured, or
+number of measurements (dark blue: 0, cyan: 1, green: 2, yellow: 3, red: 4 or
+more). a) without crystal symmetry, 2375 of 9481 reflections are measured, or
+25\%. b) with crystal symmetry, 1276 of 1427 primary reflections are measured, or
 89\%, and redundancy is also increased. }
@@
 \begin{figure}
-\caption{GUI showing details of the predicted measurement positions for a
-single $hkl$.}
+\caption{Part of a GUI showing details of the predicted measurement positions
+for a single $hkl$ and some of its equivalents.}
 \ifdograph

### 2011-02-11 docs/CrystalPlan_paper/paper.tex
Fixes #182: Final revisions for CrystalPlan paper, to be published in J Appl Cryst

@@
 \dographtrue % DO keep the graphs.
-\dographfalse % DONT put graphs - makes things faster
+%\dographfalse % DONT put graphs - makes things faster
 
@@
 CrystalPlan can help a user build an experiment plan that will acquire the most
-data possible, with sufficient coverage but limited redundancy, therefore
+unique data possible, with sufficient coverage but limited redundancy, therefore
 increasing scientific productivity. 
@@
 refinement, an experimenter will typically want to measure $ > 85\%$ of the
-peaks within a certain range of d-spacings. However, TOPAZ's wide wavelength
-band combined with its complex geometry makes it difficult to come up with
-an effective set of sample orientations that will cover all the required peaks
+unique reflections within a certain range of d-spacings. However, TOPAZ's wide
+wavelength band combined with its complex geometry makes it difficult to come up with
+an effective set of sample orientations that will cover all the required reflections
 without unnecessary redundancy.
@@
 planning their experiments. By quickly making an experimental plan that will be
-tailored to the needs of the scientist (enough peaks measured for a structure
+tailored to the needs of the scientist (enough reflections measured for a structure
 analysis), but without spending time collecting redundant data, CrystalPlan will
@@
   \item Calculate the coverage of these orientations in reciprocal space,
-  in volume and for single peaks.
+  in volume and for single reflections.
   \item Take into account crystal Bravais lattice symmetries and the higher
@@
   \item Optimize the experiment plan automatically.
-  \item Compare predicted and actual measurement of peaks.
+  \item Compare predicted and actual measurement of reflections.
 \end{itemize}
@@
 The class hierarchy allows for goniometer limitations: for example, TOPAZ is
-currently equipped with a goniometer where $\chi$ is fixed at 45 degrees.
-CrystalPlan handles this by limiting the sample orientation angles accessible to the user to $\phi$ and
-$\omega$.
+currently equipped with a goniometer where $\chi$ is fixed at 45 degrees
+(at $\omega$ to the incident beam). CrystalPlan handles this by limiting the
+sample orientation angles accessible to the user to $\phi$ and $\omega$.
 
@@
 $\gamma$) are typed in or loaded from the ISAW UB matrix file. CrystalPlan then
-generates a list of $hkl$ peaks with d-spacings larger than the $d_{min}$
-specified. For each sample orientation being simulated and for each peak, a function calculates at
+generates a list of $hkl$ reflections with d-spacings larger than the $d_{min}$
+specified. For each sample orientation being simulated and for each reflection, a function calculates at
 what wavelength $\lambda$ and in which direction it scatters. If $\lambda$ 
@@
 
-Each $hkl$ peak is a separate object that holds all predicted
+Each $hkl$ reflection is a separate object that holds all predicted
 recordings, including which detector measured it, at what (x,y) position on the
@@
 Figure \ref{fig:volume_view}a) shows a screenshot of the reciprocal space
-coverage 3D interface. The 3D view can represent {\bf Q}-space in two ways: volume coverage, or single-peak
+coverage 3D interface. The 3D view can represent {\bf Q}-space in two ways: volume coverage, or single-reflection
 coverage. The view shown can be rotated, translated or zoomed using the mouse or
@@
 Finally, both views display coverage statistics in the lower right
-corner, showing the percentage of volume (or of the number of peaks) measured,
-as well as the proportion of redundancy (voxels or peaks measured 2 or more
+corner, showing the percentage of volume (or of the number of reflections) measured,
+as well as the proportion of redundancy (voxels or reflections measured 2 or more
 times).
@@
 Figure \ref{fig:single_reflection} shows a screenshot of the GUI allowing one to
-observe in detail the predicted measurements of a particular peak, by clicking
+observe in detail the predicted measurements of a particular reflection, by clicking
 in the 3D view or by typing in the $hkl$ values. The position on the detector
-and wavelength of measurement are displayed. An additional window (not shown) allows the user to place the peak on a particular
+and wavelength of measurement are displayed. An additional window (not shown) allows the user to place the reflection on a particular
 position on a detector by automatically finding the necessary sample orientation
-angles. This is used, for example, to center important peaks or diffraction
+angles. This is used, for example, to center important reflections or diffraction
 directions in the most sensitive part of a detector.
@@
 (\ce{Na2Al2Si3O10} $\cdot$ \ce{2H2O}) single crystal was measured on a fixed
-$45\degree\ \chi$ goniometer at room temperature and ambient atmosphere. An
-initial short run (wavelength range from $\sim0.5$ \ang\ to $3.5$ \ang)
-was processed using ISAW to determine the UB matrix and cell parameters of the crystal.
+$45\degree\ \chi$ goniometer at room temperature and ambient atmosphere, with a
+wavelength range from $\sim0.5$ \ang\ to $3.5$ \ang.
+An initial short event file (each neutron detection event is saved with its
+time-of-flight and position) was processed using ISAW to determine the UB
+matrix and cell parameters of the crystal.
 
@@
 Coverage Optimizer was used to generate an experiment plan with 8 sample
-orientations (Figure \ref{fig:natrolite_gui}a) that covered 97\% of the peaks,
+orientations (Figure \ref{fig:natrolite_gui}a) that covered 97\% of the reflections,
 taking into account the crystal Laue symmetry (orthorhombic or mmm) (Figure
@@
 
-Peaks were integrated using ISAW, and the resulting .integrate file was loaded
-into CrystalPlan for comparison. The predicted and actual measurement
-positions on each detector face were compared for the 554 integrated peaks
+Reflections were integrated using ISAW, and the resulting file was
+loaded into CrystalPlan for comparison. The predicted and actual measurement
+positions on each detector face were compared for the 554 integrated reflections
 found. Good agreement was found; the root-mean-square difference between the
-predicted and actual measurement wavelength was only 0.016 \ang. The offsets in the peak
+predicted and actual measurement wavelength was only 0.016 \ang. The offsets in the reflection
 position were plotted, which helped show that the main contributor to this error
@@
 % cause?
+
+CrystalPlan was also successfully used in a similar way during experiments run
+on the TOPAZ instrument by Yuri Janssen from State University of New York, Stony
+Brook, and Henrik Clausen from University of Aarhus, Denmark.
 
@@
 diffractometers. This program will help increase scientific productivity at SNS
-beamlines such as TOPAZ.
+beamlines such as TOPAZ. With some relatively minor modifications, the program
+could also be adapted for use at reactor sources. Monochromatic reactor sources
+could be handled by adding support sweeps consisting of many motor positions;
+Laue diffractometers would requiring detecting and removing reflections with chromatic overlap.
 
@@
 recently developed capabilities for real-time data analysis to monitor
-the  signal to noise ratio on particular peaks. When the signal intensity
+the  signal to noise ratio on particular reflections. When the signal intensity
 becomes sufficient, the sample will automatically be moved to the next sample
@@
 \caption{Sample screenshot showing the 3D view of single-crystal reflection
-measurements. Each sphere represents an $hkl$ peak, color-coded to indicate the
+measurements. Each sphere represents an $hkl$ reflection, color-coded to indicate the
 number of measurements (dark blue: 0, cyan: 1, green: 2, yellow: 3, red: 4 or

### 2011-02-14 gui/gui_utils.py
Updated input to detector geometry to be the same as the translation service.

@@ -514,7 +514,7 @@ def find_parent_frame(window):
 
 
 if __name__=="__main__":
-    model.instrument.inst = model.instrument.Instrument("../instruments/TOPAZ_detectors_2010.csv")
+    model.instrument.inst = model.instrument.Instrument("../instruments/TOPAZ_geom_all_2011.csv")
     model.experiment.exp = model.experiment.Experiment(model.instrument.inst)
     save_configuration()
 #    app = wx.PySimpleApp()

### 2011-02-14 gui/panel_experiment.py

@@ -701,7 +701,7 @@ class PanelExperiment(wx.Panel):
 
 if __name__ == "__main__":
     #Ok, create the instrument
-    model.instrument.inst = model.instrument.Instrument("model/TOPAZ_detectors_all.csv")
+    model.instrument.inst = model.instrument.Instrument("model/TOPAZ_geom_all_2011.csv")
     model.instrument.inst.make_qspace()
     #Initialize the instrument and experiment
     model.experiment.exp = model.experiment.Experiment(model.instrument.inst)

### 2011-02-14 gui/panel_qspace_options.py

@@ -280,7 +280,7 @@ class PanelQspaceOptions(wx.Panel):
 
 if __name__ == '__main__':
     #Ok, create the instrument
-    model.instrument.inst = model.instrument.InstrumentInelastic("../instruments/TOPAZ_detectors_2010.csv")
+    model.instrument.inst = model.instrument.InstrumentInelastic("../instruments/TOPAZ_geom_all_2011.csv")
     model.instrument.inst.make_qspace()
     #Initialize the instrument and experiment
     model.experiment.exp = model.experiment.Experiment(model.instrument.inst)

### 2011-02-14 model/config.py

@@ -12,7 +12,7 @@ class ModelConfig(HasTraits):
 
     use_multiprocessing = Bool(False, desc='to use multiprocessing when calculating coverage.')
     reflection_divergence_degrees = Float(0.3, desc='the default half-width of the divergence of scattered beams, in degrees.')
-    default_detector_filename = String("../instruments/TOPAZ_detectors_2010.csv", desc="the default file and path to the detectors .CSV geometry specification file.")
+    default_detector_filename = String("../instruments/TOPAZ_geom_all_2011.csv", desc="the default file and path to the detectors .CSV geometry specification file.")
     force_pure_python = Bool(False, desc='to force calculations to use pure Python, rather than inline C code.')
 
 #Initialize the configuration file

### 2011-02-14 model/experiment.py

@@ -2469,7 +2469,7 @@ import copy
 class TestExperiment(unittest.TestCase):
 
     def setUp(self):
-        instrument.inst = instrument.Instrument("../instruments/TOPAZ_detectors_all.csv")
+        instrument.inst = instrument.Instrument("../instruments/TOPAZ_geom_all_2011.csv")
         instrument.inst.set_goniometer(goniometer.Goniometer() )
         self.exp = Experiment(instrument.inst)
         e = self.exp

### 2011-02-14 model/instrument.py

@@ -336,9 +336,16 @@ class Instrument:
                     det.rotation = float(row[4])
                     det.width = float(row[6]) * 10
                     det.height = float(row[7]) * 10
-                    #Calculate the pixel angles
-                    det.calculate_pixel_angles()
-                    self.detectors.append(det)
+                    # Is the detector in use?
+                    try:
+                        inuse = (float(row[8]) != 0)
+                    except:
+                        inuse = False
+                        
+                    if (inuse):
+                        #Calculate the pixel angles
+                        det.calculate_pixel_angles()
+                        self.detectors.append(det)
                 count = count + 1
         except:
             #Uh-oh! Let's go back to the original list
@@ -1672,7 +1679,7 @@ class TestInelasticInstrument(unittest.TestCase):
     """Unit test for the InstrumentInelastic class."""
     def setUp(self):
         config.cfg.force_pure_python = False
-        self.tst_inst = InstrumentInelastic("../instruments/TOPAZ_detectors_2010.csv")
+        self.tst_inst = InstrumentInelastic("../instruments/TOPAZ_geom_all_2011.csv")
         ti = self.tst_inst #@type ti InstrumentInelastic
         ti.set_goniometer(goniometer.Goniometer())
         ti.d_min = 1.0
@@ -1753,20 +1760,20 @@ class TestInstrumentWithDetectors(unittest.TestCase):
 
     def test_double_creation(self):
         assert len(self.tst_inst.detectors) == 14, "Correct # of detectors after first creation"
-        self.tst_inst = Instrument("../instruments/TOPAZ_detectors_all.csv")
-        assert len(self.tst_inst.detectors) == 48, "Correct # of detectors after second creation"
+        self.tst_inst = Instrument("../instruments/TOPAZ_geom_all_2011.csv")
+        assert len(self.tst_inst.detectors) == 14, "Correct # of detectors after second creation"
 
 
-    def test_calculate_coverage_more_detectors(self):
-        self.tst_inst = Instrument("../instruments/TOPAZ_detectors_all.csv")
-        self.tst_inst.set_goniometer(goniometer.Goniometer())
-        self.tst_inst.d_min = 0.7
-        self.tst_inst.q_resolution = 0.15
-        self.tst_inst.wl_min = 0.7
[…]

### 2011-02-14 model/optimization.py

@@ -643,7 +643,7 @@ def print_pop(ga_engine, *args):
 
 if __name__ == "__main__":
     #Inits
-    instrument.inst = instrument.Instrument("../instruments/TOPAZ_detectors_2010.csv")
+    instrument.inst = instrument.Instrument("../instruments/TOPAZ_geom_all_2011.csv")
     instrument.inst.set_goniometer(goniometer.TopazInHouseGoniometer())
     experiment.exp = experiment.Experiment(instrument.inst)
     exp = experiment.exp

### 2011-02-14 model/optimize_coverage.py

@@ -355,7 +355,7 @@ def optimize_detector_choice(optim_params, gui=False):
 def test_setup():
     """Setup a small test."""
     #Create the space
-    instrument.inst = instrument.Instrument("TOPAZ_detectors_all.csv")
+    instrument.inst = instrument.Instrument("TOPAZ_geom_all_2011.csv")
     experiment.exp = experiment.Experiment(instrument.inst)
     experiment.exp.inst.make_qspace()
     experiment.exp.verbose = False

### 2011-02-23 docs/CrystalPlan_paper/paper.tex
Crystal plan paper, revision 2 submitted to journal.

@@
 is designed to acquire single-crystal diffraction data of small-to-moderate sized unit cells with  high
-throughput. The final design calls for an array of 48 detectors covering three
-quarters of $2\pi$ steradians. However as of this writing, only 14 of the 48
-detectors are installed, giving only a quarter of $2\pi$ steradian coverage. 
-To have sufficient data for a reliable structure solution and
+throughput. The final design calls for an array of 48 detectors covering
+approximately half of $4\pi$ steradians. However as of this writing, only 14 of
+the 48 detectors are installed, giving only a sixth of $4\pi$ steradian
+coverage. To have sufficient data for a reliable structure solution and  
 refinement, an experimenter will typically want to measure $ > 85\%$ of the
 unique reflections within a certain range of d-spacings. However, TOPAZ's wide 
-wavelength band combined with its complex geometry makes it difficult to come up with
-an effective set of sample orientations that will cover all the required reflections
-without unnecessary redundancy.
+wavelength band combined with its complex geometry makes it difficult to come up
+with  an effective set of sample orientations that will cover all the required
+reflections without unnecessary redundancy.
 
@@
 beamlines such as TOPAZ. With some relatively minor modifications, the program
-could also be adapted for use at reactor sources. Monochromatic reactor sources
+could also be adapted for use at reactor sources. Monochromatic diffractometers
 could be handled by adding support sweeps consisting of many motor positions;
-Laue diffractometers would requiring detecting and removing reflections with chromatic overlap.
+Laue diffractometers on a constant intensity beam would requiring
+detecting and removing reflections with chromatic overlap.
 
@@
 from 0.9 \ang\ to 4.0\ang: a) without crystal symmetry, 17\% coverage; b) with crystal symmetry (mmm 8-fold
-symmetry), coverage increases to 65\%. }
+symmetry), coverage increases to 65\%. The area plot at the bottom shows the
+measured percentage of reciprocal volume as a function of $q$; the
+color indicates the redundancy (cyan: 1, green: 2, yellow: 3, orange: 4). }
 \ifdograph

### 2011-03-14 gui/frame_main.py
CrystalPlan: Main window a little wider.

@@ -422,7 +422,7 @@ class FrameMain(wx.Frame):
               title="%s %s - Main Window" % (CrystalPlan_version.package_name, CrystalPlan_version.version) )
         self._init_menus()
         
-        self.SetClientSize(wx.Size(800, 630))
+        self.SetClientSize(wx.Size(850, 630))
         self.SetMenuBar(self.menuBar1)
         self.SetMinSize(wx.Size(100, 100))
         self.SetAutoLayout(True)

### 2011-03-14 gui/panel_startup.py
CrystalPlan: Main window a little wider.

@@ -146,7 +146,7 @@ class StartupParameters(HasTraits):
     header_text = Group(Spring(label="Enter the parameters for Q-space simulation:", emphasized=True, show_label=True))
     second_label = Group(Spring(label="Enter the wavelength limits (from the source and/or detectors):",
                         emphasized=True, show_label=True, visible_when='show_wavelength_options'))
-    third_label = Group(Spring(label="Note: wavelength and bandwidth can also be set, on option, as a goniometer motor setting. See the goniometer tab.\nThe settings below are ignored in that case!",
+    third_label = Group(Spring(label="Note: wavelength and bandwidth can also be set, on option, as a goniometer motor setting.\n(See the goniometer tab). The settings below are ignored in that case!",
                         emphasized=False, show_label=True, visible_when='show_wavelength_options'))
 
     view = View( header_text,

### 2017-05-23 docs/CrystalPlan_paper/paper.tex
CrystalPlan: Main window a little wider.
Refs #old change CrystalPlan to OldCrystalPlan for rmp

@@
 
-\title{CrystalPlan: an Experiment Planning Tool for Crystallography}
+\title{OldCrystalPlan: an Experiment Planning Tool for Crystallography}
 
-\shorttitle{CrystalPlan: an Experiment Planning Tool}
+\shorttitle{OldCrystalPlan: an Experiment Planning Tool}
 
@@
 \aff}
-\author[]{Peter}{Peterson}
-\author[]{Xiaoping}{Wang}
+\author[]{Peter F.}{Peterson}
+\author[]{Xiaoping P.}{Wang}
 \author[]{Matthew}{Frost}
@@
 \begin{synopsis}
-Describes CrystalPlan, a software program for the automatic creation of 
+Describes OldCrystalPlan, a software program for the automatic creation of 
 crystallography experiment plans for time-of-flight diffractometers.
@@
 Beam time at large user program based x-ray and neutron scattering facilities is
-in high demand and always at a premium. CrystalPlan, a highly efficient
+in high demand and always at a premium. OldCrystalPlan, a highly efficient
 experiment planning software has been developed to maximize the use of available
@@
 reciprocal space in a single-crystal diffraction time-of-flight experiment.
-CrystalPlan can help a user build an experiment plan that will acquire the most
+OldCrystalPlan can help a user build an experiment plan that will acquire the most
 unique data possible, with sufficient coverage but limited redundancy, therefore
@@
 
-The goal of CrystalPlan was to make an easy-to-use software tool to aid users in
+The goal of OldCrystalPlan was to make an easy-to-use software tool to aid users in
 planning their experiments. By quickly making an experimental plan that will be
 tailored to the needs of the scientist (enough reflections measured for a structure
-analysis), but without spending time collecting redundant data, CrystalPlan will
+analysis), but without spending time collecting redundant data, OldCrystalPlan will
 greatly increase the scientific productivity and throughput of TOPAZ and other
@@
 
-CrystalPlan is designed with these capabilities:
+OldCrystalPlan is designed with these capabilities:
 
@@
 
-CrystalPlan currently supports area detectors with flat faces and arbitrary
+OldCrystalPlan currently supports area detectors with flat faces and arbitrary
 rectangular dimensions; this reflects the detector types in use at SNS (Anger
@@
 
-CrystalPlan can accomodate a variety of sample
+OldCrystalPlan can accomodate a variety of sample
 orientation goniometers. A base Goniometer class (in the object-oriented
@@
 currently equipped with a goniometer where $\chi$ is fixed at 45 degrees
-(at $\omega$ to the incident beam). CrystalPlan handles this by limiting the
+(at $\omega$ to the incident beam). OldCrystalPlan handles this by limiting the
 sample orientation angles accessible to the user to $\phi$ and $\omega$.
@@
 text file along with the lattice parameters. 
-CrystalPlan can load this UB matrix file and uses it for reciprocal coverage
+OldCrystalPlan can load this UB matrix file and uses it for reciprocal coverage
 calculations and optimization.
@@
 Crystal lattice parameters (lengths $a$, $b$, $c$ and angles $\alpha$, $\beta$,
-$\gamma$) are typed in or loaded from the ISAW UB matrix file. CrystalPlan then
+$\gamma$) are typed in or loaded from the ISAW UB matrix file. OldCrystalPlan then
 generates a list of $hkl$ reflections with d-spacings larger than the $d_{min}$
@@
 entries. Each entry in the $4^{th}$ dimension points to a voxel that is an
-equivalent due to crystal symmetry. This allows CrystalPlan to quickly compute a volume coverage
+equivalent due to crystal symmetry. This allows OldCrystalPlan to quickly compute a volume coverage
 map that takes into account crystal symmetry and the resulting redundancy: at
@@
 
-CrystalPlan allows the user to manually enter lists of sample orientations to
+OldCrystalPlan allows the user to manually enter lists of sample orientations to
 simulate and graphically display the coverage in reciprocal space;
@@
 time consuming and only feasible for an experienced crytallographer.
-To make the process more efficient, CrystalPlan includes an Automatic
+To make the process more efficient, OldCrystalPlan includes an Automatic
 Coverage Optimizer, which takes as its inputs the desired number of sample
@@
 
-To this end, CrystalPlan was written in Python \cite{python}, a powerful, open,
+To this end, OldCrystalPlan was written in Python \cite{python}, a powerful, open,
 cross-platform scripting language. Python was chosen for its ease of cross-platform deployment, and for
@@
 
-The calculation code of CrystalPlan is placed in a model layer and kept
+The calculation code of OldCrystalPlan is placed in a model layer and kept
 completely separate from eventual GUI elements, using a Model/View-Controller
@@
 An attractive and easy-to-use GUI was considered an important part of
-CrystalPlan, since the program is meant to be used interactively as a guide to
+OldCrystalPlan, since the program is meant to be used interactively as a guide to
 the experimenter whether new or experienced. This requires intuitive GUI
@@
 
-The resulting UB matrix was loaded into CrystalPlan, where the Automatic
+The resulting UB matrix was loaded into OldCrystalPlan, where the Automatic
 Coverage Optimizer was used to generate an experiment plan with 8 sample
@@
 Reflections were integrated using ISAW, and the resulting file was
-loaded into CrystalPlan for comparison. The predicted and actual measurement
+loaded into OldCrystalPlan for comparison. The predicted and actual measurement
 positions on each detector face were compared for the 554 integrated reflections
@@
 
-CrystalPlan was also successfully used in a similar way during experiments run
+OldCrystalPlan was also successfully used in a similar way during experiments run
 on the TOPAZ instrument by Yuri Janssen from State University of New York, Stony
@@
 
-We have described CrystalPlan, a software program designed to quickly produce
+We have described OldCrystalPlan, a software program designed to quickly produce
 efficient experiment plans for wide bandwidth time-of-flight Laue
@@
 
-One planned future development for CrystalPlan include extending the software
+One planned future development for OldCrystalPlan include extending the software
 to help predict coverage in inelastic neutron scattering experiments. Another
@@
 right time to end data acquisition at a particular sample orientation.
-CrystalPlan would interface with data analysis software such as ISAW and use
+OldCrystalPlan would interface with data analysis software such as ISAW and use
 recently developed capabilities for real-time data analysis to monitor

### 2017-05-23 OldCrystalPlan_version.py

@@ -1,4 +1,4 @@
-"""Version information for CrystalPlan"""
+"""Version information for OldCrystalPlan"""
 # Author: Janik Zikovsky, zikovskyjl@ornl.gov
 # Version: $Id$
 
@@ -17,9 +17,9 @@
 #    return revision
 
 version = '1.2'
-package_name = 'CrystalPlan'
+package_name = 'OldCrystalPlan'
 description = \
-"""CrystalPlan is an experiment planning tool for crystallography. You can choose an instrument and supply your sample's lattice parameters to simulate which reflections will be measured, by which detectors and at what wavelengths."""
+"""OldCrystalPlan is an experiment planning tool for crystallography. You can choose an instrument and supply your sample's lattice parameters to simulate which reflections will be measured, by which detectors and at what wavelengths."""
 
 license = """Enter license text here."""
 author = "Janik Zikovsky"
@@ -28,10 +28,10 @@ url = 'http://neutronsr.us'
 copyright = '(C) 2010'
 
 #Path to icons, relative to gui scripts path
-icon_file = "icons/CrystalPlan_icon.png"
-icon_file_config = "icons/CrystalPlan_icon_config.png"
-icon_file_3d = "icons/CrystalPlan_icon_3d.png"
-icon_file_optimizer = "icons/CrystalPlan_icon_optimizer.png"
+icon_file = "icons/OldCrystalPlan_icon.png"
+icon_file_config = "icons/OldCrystalPlan_icon_config.png"
+icon_file_3d = "icons/OldCrystalPlan_icon_3d.png"
+icon_file_optimizer = "icons/OldCrystalPlan_icon_optimizer.png"
 
 if __name__ == "__main__":
     print package_name, version

### 2017-05-23 __init__.py

@@ -1,11 +1,11 @@
-"""CrystalPlan package file."""
+"""OldCrystalPlan package file."""
 # Author: Janik Zikovsky, zikovskyjl@ornl.gov
-# Version: $Id: CrystalPlan.py 1127 2010-04-01 19:28:43Z 8oz $
+# Version: $Id: OldCrystalPlan.py 1127 2010-04-01 19:28:43Z 8oz $
 
-#print "CrystalPlan module being imported from", __file__
-#print "CrystalPlan module, __name__ is", __name__
+#print "OldCrystalPlan module being imported from", __file__
+#print "OldCrystalPlan module, __name__ is", __name__
 
-from CrystalPlan_version import version as __version__
+from OldCrystalPlan_version import version as __version__
 
 #Manipulate the PYTHONPATH to put model directly in view of it
 #   This way, "import model" works.
@@ -20,4 +20,4 @@ import gui
 import pyevolve
 
 #And the stuff in here
-import CrystalPlan_version
+import OldCrystalPlan_version

### 2017-05-23 doc_maker/__init__.py

@@ -1,4 +1,4 @@
-"""__init__.py module setup for the CrystalPlan GUI documentation maker (doc_maker) code."""
+"""__init__.py module setup for the OldCrystalPlan GUI documentation maker (doc_maker) code."""
 # Author: Janik Zikovsky, zikovskyjl@ornl.gov
 # Version: $Id: __init__.py 1127 2010-04-01 19:28:43Z 8oz $
 

### 2017-05-23 doc_maker/user_guide.py

@@ -1,5 +1,5 @@
 """Module to generate screenshots for the
-CrystalPlan user guide.
+OldCrystalPlan user guide.
 """
 
 import wx
@@ -706,7 +706,7 @@ class FakeClickObject:
 
 #========================================================================================================
 def generate_user_guide(fm, fv):
-    """Script to generate screenshots for the CrystalPlan user guide.
+    """Script to generate screenshots for the OldCrystalPlan user guide.
 
     Parameters:
         fm: the FrameMain object.
@@ -714,7 +714,7 @@ def generate_user_guide(fm, fv):
     """
     #Make sure we don't load the .pyc file
     filename = os.path.splitext(__file__)[0] + ".py"
-#    filename = "/home/janik/Code/GenUtils/trunk/python/CrystalPlan/doc_maker/user_guide.py"
+#    filename = "/home/janik/Code/GenUtils/trunk/python/OldCrystalPlan/doc_maker/user_guide.py"
     print "Reading script from", filename
     #Read the code
     code = []

### 2017-05-23 gui/__init__.py

@@ -1,8 +1,8 @@
-"""__init__.py module setup for the CrystalPlan GUI code."""
+"""__init__.py module setup for the OldCrystalPlan GUI code."""
 # Author: Janik Zikovsky, zikovskyjl@ornl.gov
 # Version: $Id: __init__.py 1127 2010-04-01 19:28:43Z 8oz $
 
-#print "CrystalPlan.gui is being imported from ", __file__
+#print "OldCrystalPlan.gui is being imported from ", __file__
 
 #We import the most important modules only.
 import frame_main

### 2017-05-23 gui/dialog_preferences.py

@@ -6,7 +6,7 @@
 
 #--- General Imports ---
 import wx
-import CrystalPlan_version
+import OldCrystalPlan_version
 
 #--- GUI Imports ---
 import config_gui
@@ -68,7 +68,7 @@ class DialogPreferences(wx.Dialog):
         self.SetClientSize(wx.Size(500, 400))
 
         import os
-        icon_file = os.path.join(os.path.dirname(__file__), CrystalPlan_version.icon_file_config)
+        icon_file = os.path.join(os.path.dirname(__file__), OldCrystalPlan_version.icon_file_config)
         self.SetIcon( wx.Icon(icon_file, wx.BITMAP_TYPE_PNG) )
 
         self.buttonOK = wx.Button(id=wxID_DIALOGPREFERENCESBUTTONOK,

### 2017-05-23 gui/frame_main.py

@@ -1,4 +1,4 @@
-"""FrameMain: main GUI window for the CrystalPlan application.
+"""FrameMain: main GUI window for the OldCrystalPlan application.
 """
 import os.path
 # Author: Janik Zikovsky, zikovskyjl@ornl.gov
@@ -22,7 +22,7 @@ import gui_utils
 import frame_qspace_view
 import frame_reflection_info
 import dialog_preferences
-import CrystalPlan_version
+import OldCrystalPlan_version
 import display_thread
 import frame_optimizer
 import detector_plot
@@ -152,7 +152,7 @@ class FrameMain(wx.Frame):
         self.Bind(wx.EVT_MENU, self.OnMenuUserGuide, id=id)
 
         id = wx.NewId()
-        parent.Append(id=id, help='', kind=wx.ITEM_NORMAL, text=u'About %s...' % CrystalPlan_version.package_name)
+        parent.Append(id=id, help='', kind=wx.ITEM_NORMAL, text=u'About %s...' % OldCrystalPlan_version.package_name)
         self.Bind(wx.EVT_MENU, self.OnMenuAbout, id=id)
 
 #        parent.AppendSeparator()
@@ -343,20 +343,20 @@ class FrameMain(wx.Frame):
 
     def OnMenuAbout(self, event):
         info = wx.AboutDialogInfo()
-        info.SetName(CrystalPlan_version.package_name)
-        info.SetVersion(CrystalPlan_version.version)
-        info.SetDescription(CrystalPlan_version.description)
-        info.SetCopyright(CrystalPlan_version.copyright)
-        info.AddDeveloper(CrystalPlan_version.author + " (" + CrystalPlan_version.author_email + ")")
-        info.AddDocWriter(CrystalPlan_version.author + " (" + CrystalPlan_version.author_email + ")")
+        info.SetName(OldCrystalPlan_version.package_name)
+        info.SetVersion(OldCrystalPlan_version.version)
+        info.SetDescription(OldCrystalPlan_version.description)
+        info.SetCopyright(OldCrystalPlan_version.copyright)
+        info.AddDeveloper(OldCrystalPlan_version.author + " (" + OldCrystalPlan_version.author_email + ")")
+        info.AddDocWriter(OldCrystalPlan_version.author + " (" + OldCrystalPlan_version.author_email + ")")
[…]

### 2017-05-23 gui/frame_optimizer.py

@@ -22,7 +22,7 @@ import gui_utils
 #--- Model Imports ----
 import model
 from model.optimization import OptimizationParameters
-import CrystalPlan_version
+import OldCrystalPlan_version
 
 if __name__=="__main__":
     #Manipulate the PYTHONPATH to put model directly in view of it
@@ -537,7 +537,7 @@ class FrameOptimizer(wx.Frame):
 
         #Little icon
         import os
-        icon_file = os.path.join(os.path.dirname(__file__), CrystalPlan_version.icon_file_optimizer)
+        icon_file = os.path.join(os.path.dirname(__file__), OldCrystalPlan_version.icon_file_optimizer)
         self.SetIcon( wx.Icon(icon_file, wx.BITMAP_TYPE_PNG) )
 
         self.splitterMain = wx.SplitterWindow(id=wxID_FRAMEOPTIMIZERSPLITTERMAIN,

### 2017-05-23 gui/frame_qspace_view.py

@@ -20,7 +20,7 @@ import display_thread
 #--- Model Imports ---
 import model
 from model.numpy_utils import column, vector_length
-import CrystalPlan_version
+import OldCrystalPlan_version
 
 #--- Mayavi Imports ---
 try:
@@ -934,7 +934,7 @@ class FrameQspaceView(wx.Frame):
         self._init_ctrls(parent)
         #Set the icon
         import os
-        icon_file = os.path.join(os.path.dirname(__file__), CrystalPlan_version.icon_file_3d)
+        icon_file = os.path.join(os.path.dirname(__file__), OldCrystalPlan_version.icon_file_3d)
         self.SetIcon( wx.Icon(icon_file, wx.BITMAP_TYPE_PNG) )
 
         #Create the qspace options panel

### 2017-05-23 gui/gui_utils.py

@@ -113,7 +113,7 @@ def dialog_to_save_experiment_to_CSV(parent):
 last_experiment_path = ''
 def save_experiment_file_dialog(parent):
     """Opens a dialog asking the user where to save the experiment."""
-    filters = 'CrystalPlan Experiment files (*.exp)|*.exp|All files (*)|*|'
+    filters = 'OldCrystalPlan Experiment files (*.exp)|*.exp|All files (*)|*|'
     if is_mac(): filters = '' #Filters tend to crash on mac
     global last_experiment_path
     (path, filename) = os.path.split(last_experiment_path)
@@ -132,7 +132,7 @@ def save_experiment_file_dialog(parent):
 
 def load_experiment_file_dialog(parent):
     """Opens a dialog asking the user where to load the experiment."""
-    filters = 'CrystalPlan Experiment files (*.exp)|*.exp|All files (*)|*|'
+    filters = 'OldCrystalPlan Experiment files (*.exp)|*.exp|All files (*)|*|'
     if is_mac(): filters = '' #Filters tend to crash on mac
     global last_experiment_path
     (path, filename) = os.path.split(last_experiment_path)

### 2017-05-23 gui/main.py

@@ -1,10 +1,10 @@
 #!/usr/bin/env python
-"""CrystalPlan GUI Main file.
+"""OldCrystalPlan GUI Main file.
 Includes GUI launcher, logger, error handler."""
-#Boa:App:CrystalPlanApp
+#Boa:App:OldCrystalPlanApp
 
 # Author: Janik Zikovsky, zikovskyjl@ornl.gov
-# Version: $Id: CrystalPlan.py 1127 2010-04-01 19:28:43Z 8oz $
+# Version: $Id: OldCrystalPlan.py 1127 2010-04-01 19:28:43Z 8oz $
 
 #--- General Imports ---
 from __future__ import with_statement
@@ -109,7 +109,7 @@ def excepthook(type, value, tb, thread_information="Main Loop"):
 
 
 #-------------------------------------------------------------------------
-class CrystalPlanApp(wx.App):
+class OldCrystalPlanApp(wx.App):
     def OnInit(self):
         #Create the main GUI frame
         import frame_main
@@ -125,17 +125,17 @@ class CrystalPlanApp(wx.App):
 
 #-------------------------------------------------------------------------
 def launch_gui(inelastic, hb3a):
-    """Launch the CrystalPlan GUI.
+    """Launch the OldCrystalPlan GUI.
 
     Parameters:
         inelastic: boolean, to indicate whether the instrument is for inelastic scattering.
         hb3a: boolean, for the HB3A beamline
         """
 
-    import CrystalPlan_version
+    import OldCrystalPlan_version
         
     #Since imports take a while, print out this status line first.
-    print "-------------- %s %s GUI is starting -----------------" % (CrystalPlan_version.package_name, CrystalPlan_version.version)
+    print "-------------- %s %s GUI is starting -----------------" % (OldCrystalPlan_version.package_name, OldCrystalPlan_version.version)
     
     #--- GUI Imports ---
     import display_thread
@@ -145,7 +145,7 @@ def launch_gui(inelastic, hb3a):
 
     #Create a StdOut wrapper
     global out_wrapper
-    out_wrapper = OutWrapper(sys.stdout, r'/tmp/CrystalPlan_Log_' + datetime.datetime.now().strftime("%Y-%m-%d_%H-%M-%S") + ".txt" )
[…]

### 2017-05-23 gui/panel_startup.py

@@ -269,7 +269,7 @@ class PanelStartup(wx.Panel):
         self.SetClientSize(wx.Size(600, 600))
 
         self.staticTextHelp = wx.StaticText(id=wxID_DIALOGSTARTUPSTATICTEXTHELP,
-              label=u"The CrystalPlan application is used to simulate the coverage of reciprocal space of an instrument's detectors, given a list of sample orientations.",
+              label=u"The OldCrystalPlan application is used to simulate the coverage of reciprocal space of an instrument's detectors, given a list of sample orientations.",
               name=u'staticTextHelp', parent=self, pos=wx.Point(0, 40),
               style=0)
         if gui_utils.is_mac():

### 2017-05-23 model/__init__.py

@@ -2,8 +2,8 @@
 # Author: Janik Zikovsky, zikovskyjl@ornl.gov
 # Version: $Id$
 
-#print "CrystalPlan.model being imported from", __file__
-#print "CrystalPlan.model, __name__ is", __name__
+#print "OldCrystalPlan.model being imported from", __file__
+#print "OldCrystalPlan.model, __name__ is", __name__
 
 #We import all the modules in this so as to make "import model" do all the necessary work.
 import config

### 2017-05-23 model/crystals.py

@@ -286,7 +286,7 @@ class Crystal(HasTraits):
         
         - Start with the a* vector along the beam axis
            = in the same direction as the neutron movement
-           = +Z in our usual convention (e.g. in ISAW, CrystalPlan and Mantid)
+           = +Z in our usual convention (e.g. in ISAW, OldCrystalPlan and Mantid)
            = +X axis in the MOSFLM convention
         - Start with the c* vector pointing towards the rotation axis
            = the vertical axis pointing upwards (+Y in our usual convention).

### 2017-05-23 model/tools.py

@@ -1,4 +1,4 @@
-"""Various tools and other CrystalPlan useful calculations.
+"""Various tools and other OldCrystalPlan useful calculations.
 """
 
 # Author: Janik Zikovsky, zikovskyjl@ornl.gov

### 2017-05-23 oldcrystalplan.py

@@ -1,7 +1,7 @@
 #!/usr/bin/env python
-"""The CrystalPlan application.
+"""The OldCrystalPlan application.
 
-CrystalPlan is an experiment planning tool for crystallography.
+OldCrystalPlan is an experiment planning tool for crystallography.
 You can choose an instrument and supply your sample's lattice
 parameters to simulate which reflections will be measured,
 by which detectors and at what wavelengths.
@@ -13,5 +13,5 @@ Version: $Id$
 # Version: $Id$
 
 #Simply import and launch the GUI
-import CrystalPlan.gui.main
-CrystalPlan.gui.main.handle_arguments_and_launch(InstalledVersion=True)
+import OldCrystalPlan.gui.main
+OldCrystalPlan.gui.main.handle_arguments_and_launch(InstalledVersion=True)

### 2017-05-23 setup.py

@@ -1,5 +1,5 @@
 #!/usr/bin/env python
-"""Script for installing the CrystalPlan utility."""
+"""Script for installing the OldCrystalPlan utility."""
 
 # Author: Janik Zikovsky, zikovskyjl@ornl.gov
 # Version: $Id$
@@ -7,20 +7,20 @@
 #--- Imports ---
 from setuptools import setup, find_packages
 import sys
-import CrystalPlan_version
+import OldCrystalPlan_version
 
 #Two packages: the GUI and the model code
 packages = find_packages()
-packages = ['CrystalPlan', 'CrystalPlan.model',  'CrystalPlan.pyevolve', 'CrystalPlan.gui', 'CrystalPlan.model.pygene']
-package_dir = {'CrystalPlan': '.',  'CrystalPlan.pyevolve':'pyevolve', 'CrystalPlan.model':'model', 'CrystalPlan.gui':'gui', 'CrystalPlan.model.pygene':'model/pygene'}
+packages = ['OldCrystalPlan', 'OldCrystalPlan.model',  'OldCrystalPlan.pyevolve', 'OldCrystalPlan.gui', 'OldCrystalPlan.model.pygene']
+package_dir = {'OldCrystalPlan': '.',  'OldCrystalPlan.pyevolve':'pyevolve', 'OldCrystalPlan.model':'model', 'OldCrystalPlan.gui':'gui', 'OldCrystalPlan.model.pygene':'model/pygene'}
 #data_files = [ ('instruments', './instruments/*.csv'), ('instruments', './instruments/*.xls') ]
 data_files = []
-package_data = {'CrystalPlan':['instruments/*.xls', 'instruments/*.csv', 'instruments/*.detcal',
+package_data = {'OldCrystalPlan':['instruments/*.xls', 'instruments/*.csv', 'instruments/*.detcal',
                                'docs/*.*', 'docs/animations/*.*', 'docs/eq/*.*', 'docs/screenshots/*.*' ],
-    'CrystalPlan.model':['data/*.*'],
-    'CrystalPlan.gui':['icons/*.png']
+    'OldCrystalPlan.model':['data/*.*'],
+    'OldCrystalPlan.gui':['icons/*.png']
 }
-scripts = ['crystalplan.py']
+scripts = ['oldcrystalplan.py']
 
 #Package requirements
[…]
